### 2014-09-10 RFCs/HDF5/VOL/user_guide/new_api.tex

\section{New VOL API Routines}
New API routines have been added to the HDF5 library to manage VOL plugins. This section details each new API call and explains their intended usage.

\begin{lstlisting}
  hid_t H5VLobject_register(void *obj, H5I_type_t obj_type, const H5VL_class_t *cls);
\end{lstlisting}
to register an {\tt hid\_t} with an object {\tt obj} associated with
the VOL plugin of class {\tt cls}. This is needed in all iterate and
visit callbacks where the plugins internally need to wrap an {\tt
  hid\_t} around an object to call the user defined operation {\tt op}
on.

\begin{lstlisting}
  herr_t H5VLget_object(hid_t obj_id, void **obj, H5VL_t **vol_plugin);
\end{lstlisting}
to retrieve the VOL object and plugin structure from an HDF5
identifier ({\tt hid\_t}). The plugin structure is defined as:
\begin{lstlisting}
struct H5VL_t {
    const H5VL_class_t *cls;     /* constant class info */
    const char *container_name;  /* name of the underlying storage 
                                    container */
    unsigned long feature_flags; /* VOL Driver feature Flags */
    int nrefs;                   /* number of references by objects
                                    using this struct */
};
\end{lstlisting}

% H5_DLL void *H5Pget_vol_info(hid_t plist_id);
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: