### 2014-09-23 RFCs/HDF5/VOL/user_guide/appendix.tex
add example
Former-commit-id: 06d8b556668e3dd53df67be09620df7dc29f6629

\appendix
\section{VOL Plugin Example}
\label{sec:A}

This appendix shows a simple implementation of a {\tt printf} plugin that stacks on top of the native HDF5 plugin. Note that not all callbacks have been implemented since this is just an example for explaining the general framework of the VOL plugin.

\begin{lstlisting}

#include <stdio.h>
#include <stdlib.h>
#include <assert.h>
#include "hdf5.h"

#define LOG 502

static herr_t H5VL_log_init(hid_t vipl_id);
static herr_t H5VL_log_term(hid_t vtpl_id);

/* Datatype callbacks */
static void *H5VL_log_datatype_commit(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id, hid_t dxpl_id, void **req);
static void *H5VL_log_datatype_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t tapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_datatype_get(void *dt, H5VL_datatype_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
static herr_t H5VL_log_datatype_close(void *dt, hid_t dxpl_id, void **req);

/* Dataset callbacks */
static void *H5VL_log_dataset_create(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req);
static void *H5VL_log_dataset_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_dataset_read(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                                    hid_t file_space_id, hid_t plist_id, void *buf, void **req);
static herr_t H5VL_log_dataset_write(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                                     hid_t file_space_id, hid_t plist_id, const void *buf, void **req);
static herr_t H5VL_log_dataset_close(void *dset, hid_t dxpl_id, void **req);

/* File callbacks */
static void *H5VL_log_file_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id, void **req);
static void *H5VL_log_file_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_file_get(void *file, H5VL_file_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
static herr_t H5VL_log_file_close(void *file, hid_t dxpl_id, void **req);

/* Group callbacks */
static void *H5VL_log_group_create(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_group_close(void *grp, hid_t dxpl_id, void **req);

/* Link callbacks */

/* Object callbacks */
static void *H5VL_log_object_open(void *obj, H5VL_loc_params_t loc_params, H5I_type_t *opened_type, hid_t dxpl_id, void **req);
static herr_t H5VL_log_object_specific(void *obj, H5VL_loc_params_t loc_params, H5VL_object_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);

hid_t native_plugin_id = -1;

static const H5VL_class_t H5VL_log_g = {
    0,
    LOG,
    "log",					/* name */
    H5VL_log_init,                              /* initialize */
    H5VL_log_term,                              /* terminate */
    sizeof(hid_t),
    NULL,
    NULL,
    {                                           /* attribute_cls */
        NULL, //H5VL_log_attr_create,                /* create */
        NULL, //H5VL_log_attr_open,                  /* open */
        NULL, //H5VL_log_attr_read,                  /* read */
        NULL, //H5VL_log_attr_write,                 /* write */
        NULL, //H5VL_log_attr_get,                   /* get */
        NULL, //H5VL_log_attr_specific,              /* specific */
        NULL, //H5VL_log_attr_optional,              /* optional */
        NULL  //H5VL_log_attr_close                  /* close */
    },
    {                                           /* dataset_cls */
        H5VL_log_dataset_create,                    /* create */
        H5VL_log_dataset_open,                      /* open */
        H5VL_log_dataset_read,                      /* read */
        H5VL_log_dataset_write,                     /* write */
        NULL, //H5VL_log_dataset_get,               /* get */
        NULL, //H5VL_log_dataset_specific,          /* specific */
        NULL, //H5VL_log_dataset_optional,          /* optional */
        H5VL_log_dataset_close                      /* close */
    },
    {                                               /* datatype_cls */
        H5VL_log_datatype_commit,                   /* commit */
        H5VL_log_datatype_open,                     /* open */
        H5VL_log_datatype_get,                      /* get_size */
        NULL, //H5VL_log_datatype_specific,         /* specific */
        NULL, //H5VL_log_datatype_optional,         /* optional */
        H5VL_log_datatype_close                     /* close */
    },
    {                                           /* file_cls */
        H5VL_log_file_create,                      /* create */
        H5VL_log_file_open,                        /* open */
        H5VL_log_file_get,                         /* get */
        NULL, //H5VL_log_file_specific,            /* specific */
        NULL, //H5VL_log_file_optional,            /* optional */
        H5VL_log_file_close                        /* close */
    },
    {                                           /* group_cls */
        H5VL_log_group_create,                     /* create */
        NULL, //H5VL_log_group_open,               /* open */
        NULL, //H5VL_log_group_get,                /* get */
        NULL, //H5VL_log_group_specific,           /* specific */
        NULL, //H5VL_log_group_optional,           /* optional */
        H5VL_log_group_close                       /* close */
    },
    {                                           /* link_cls */
        NULL, //H5VL_log_link_create,                /* create */
        NULL, //H5VL_log_link_copy,                  /* copy */
        NULL, //H5VL_log_link_move,                  /* move */
        NULL, //H5VL_log_link_get,                   /* get */
        NULL, //H5VL_log_link_specific,              /* specific */
        NULL, //H5VL_log_link_optional,              /* optional */
    },
    {                                           /* object_cls */
        H5VL_log_object_open,                        /* open */
        NULL, //H5VL_log_object_copy,                /* copy */
        NULL, //H5VL_log_object_get,                 /* get */
        H5VL_log_object_specific,                    /* specific */
        NULL, //H5VL_log_object_optional,            /* optional */
    },
    {
        NULL,
        NULL,
        NULL
    },
    NULL
};

typedef struct H5VL_log_t {
    void   *under_object;
} H5VL_log_t;

static herr_t
visit_cb(hid_t oid, const char *name,
         const H5O_info_t *oinfo, void *udata)
{
    ssize_t len;
    char n[25];

    if(H5Iget_type(oid) == H5I_GROUP) {
        len = H5VLget_plugin_name(oid, n, 50);
        printf ("Visiting GROUP VOL name = %s  %d\n", n, len);
    }
    if(H5Iget_type(oid) == H5I_DATASET) 
        printf("visiting dataset\n");
    if(H5Iget_type(oid) == H5I_DATATYPE) 
        printf("visiting datatype\n");

    return 1;
} /* end h5_verify_cached_stabs_cb() */

int main(int argc, char **argv) {
        const char file_name[]="large_dataset.h5";
	const char group_name[]="/Group";
	const char dataset_name[]="Data";
	char fullpath[500];
	hid_t file_id;
	hid_t group_id;
	hid_t dataspaceId;
        hid_t datasetId;
        hid_t acc_tpl;
        hid_t under_fapl;
        hid_t vol_id, vol_id2;
        hid_t int_id;
        hid_t attr;
        hid_t space;
	const unsigned int nelem=60;
	int *data;
	unsigned int i;
	hsize_t dims[1];
        ssize_t len;
        char name[25];
        static hsize_t      ds_size[2] = {10, 20};

        under_fapl = H5Pcreate (H5P_FILE_ACCESS);
        H5Pset_fapl_native(under_fapl);
        assert(H5VLis_registered("native") == 1);

        vol_id = H5VLregister (&H5VL_log_g);
        assert(vol_id > 0);
        assert(H5VLis_registered("log") == 1);

        vol_id2 = H5VLget_plugin_id("log");
        H5VLinitialize(vol_id2, H5P_DEFAULT);
        H5VLclose(vol_id2);

        native_plugin_id = H5VLget_plugin_id("native");
        assert(native_plugin_id > 0);

        acc_tpl = H5Pcreate (H5P_FILE_ACCESS);
        H5Pset_vol(acc_tpl, vol_id, &under_fapl);

	file_id = H5Fcreate(file_name, H5F_ACC_TRUNC, H5P_DEFAULT, acc_tpl);
        len = H5VLget_plugin_name(file_id, name, 25);
        printf ("FILE VOL name = %s  %d\n", name, len);

	group_id = H5Gcreate2(file_id, group_name, H5P_DEFAULT, H5P_DEFAULT, H5P_DEFAULT);
        len = H5VLget_plugin_name(group_id, name, 50);
        printf ("GROUP VOL name = %s  %d\n", name, len);

        int_id = H5Tcopy(H5T_NATIVE_INT);
        H5Tcommit2(file_id, "int", int_id, H5P_DEFAULT, H5P_DEFAULT, H5P_DEFAULT);
        len = H5VLget_plugin_name(int_id, name, 50);
        printf ("DT COMMIT name = %s  %d\n", name, len);
        H5Tclose(int_id);

        int_id = H5Topen2(file_id, "int", H5P_DEFAULT);
        len = H5VLget_plugin_name(int_id, name, 50);
        printf ("DT OPEN name = %s  %d\n", name, len);
        H5Tclose(int_id);

        int_id = H5Oopen(file_id,"int",H5P_DEFAULT);
        len = H5VLget_plugin_name(int_id, name, 50);
        printf ("DT OOPEN name = %s  %d\n", name, len);

        len = H5Fget_name(file_id, name, 50);
        printf("name = %d  %s\n", len, name);

	data = malloc (sizeof(int)*nelem);
	for(i=0;i<nelem;++i)
	  data[i]=i;

	dims [0] = 60;
	dataspaceId = H5Screate_simple(1, dims, NULL); 
        space = H5Screate_simple (2, ds_size, ds_size);

	sprintf(fullpath,"%s/%s",group_name,dataset_name);
	datasetId = H5Dcreate2(file_id, fullpath, H5T_NATIVE_INT, dataspaceId, H5P_DEFAULT, H5P_DEFAULT, H5P_DEFAULT);
	H5Sclose(dataspaceId);

        len = H5VLget_plugin_name(datasetId, name, 50);
        printf ("DSET name = %s  %d\n", name, len);

	H5Dwrite(datasetId, H5T_NATIVE_INT, H5S_ALL, H5S_ALL, H5P_DEFAULT, data);
	H5Dclose(datasetId);

        H5Ovisit(file_id, H5_INDEX_NAME, H5_ITER_NATIVE, visit_cb, NULL);

	free (data);
        H5Oclose(int_id);
        H5Sclose (space);
	H5Gclose(group_id);

	H5Fclose(file_id);
        H5Pclose(acc_tpl);
        H5Pclose(under_fapl);

        H5VLclose(native_plugin_id);
        H5VLterminate(vol_id, H5P_DEFAULT);
        H5VLunregister (vol_id);
        assert(H5VLis_registered("log") == 0);
	return 0;
}

static herr_t H5VL_log_init(hid_t vipl_id)
{
    printf("------- LOG INIT\n");
    return 0;
}

static herr_t H5VL_log_term(hid_t vtpl_id)
{
    printf("------- LOG TERM\n");
    return 0;
}

static void *
H5VL_log_file_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id, void **req)
{
    hid_t under_fapl;
    H5VL_log_t *file;

    file = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    under_fapl = *((hid_t *)H5Pget_vol_info(fapl_id));
    file->under_object = H5VLfile_create(name, flags, fcpl_id, under_fapl, dxpl_id, req);

    printf("------- LOG H5Fcreate\n");
    return (void *)file;
}

static void *
H5VL_log_file_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req)
{
    hid_t under_fapl;
    H5VL_log_t *file;

    file = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    under_fapl = *((hid_t *)H5Pget_vol_info(fapl_id));
    file->under_object = H5VLfile_open(name, flags, under_fapl, dxpl_id, req);

    printf("------- LOG H5Fopen\n");
    return (void *)file;
}

static herr_t 
H5VL_log_file_get(void *file, H5VL_file_get_t get_type, hid_t dxpl_id, void **req, va_list arguments)
{
    H5VL_log_t *f = (H5VL_log_t *)file;

    H5VLfile_get(f->under_object, native_plugin_id, get_type, dxpl_id, req, arguments);

    printf("------- LOG H5Fget %d\n", get_type);
    return 1;
}
static herr_t 
H5VL_log_file_close(void *file, hid_t dxpl_id, void **req)
{
    H5VL_log_t *f = (H5VL_log_t *)file;

    H5VLfile_close(f->under_object, native_plugin_id, dxpl_id, req);
    free(f);

    printf("------- LOG H5Fclose\n");
    return 1;
}

static void *
H5VL_log_group_create(void *obj, H5VL_loc_params_t loc_params, const char *name, 
                      hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *group;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    group = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    group->under_object = H5VLgroup_create(o->under_object, loc_params, native_plugin_id, name, gcpl_id,  gapl_id, dxpl_id, req);

    printf("------- LOG H5Gcreate\n");
    return (void *)group;
}

static herr_t 
H5VL_log_group_close(void *grp, hid_t dxpl_id, void **req)
{
    H5VL_log_t *g = (H5VL_log_t *)grp;

    H5VLgroup_close(g->under_object, native_plugin_id, dxpl_id, req);
    free(g);

    printf("------- LOG H5Gclose\n");
    return 1;
}

static void *
H5VL_log_datatype_commit(void *obj, H5VL_loc_params_t loc_params, const char *name, 
                         hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *dt;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dt = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dt->under_object = H5VLdatatype_commit(o->under_object, loc_params, native_plugin_id, name, 
                                           type_id, lcpl_id, tcpl_id, tapl_id, dxpl_id, req);

    printf("------- LOG H5Tcommit\n");
    return dt;
}
static void *
H5VL_log_datatype_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t tapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *dt;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dt = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dt->under_object = H5VLdatatype_open(o->under_object, loc_params, native_plugin_id, name, tapl_id, dxpl_id, req);

    printf("------- LOG H5Topen\n");
    return (void *)dt;
}

static herr_t 
H5VL_log_datatype_get(void *dt, H5VL_datatype_get_t get_type, hid_t dxpl_id, void **req, va_list arguments)
{
    H5VL_log_t *o = (H5VL_log_t *)dt;
    herr_t ret_value;

    ret_value = H5VLdatatype_get(o->under_object, native_plugin_id, get_type, dxpl_id, req, arguments);

    printf("------- LOG datatype get\n");
    return ret_value;
}

static herr_t 
H5VL_log_datatype_close(void *dt, hid_t dxpl_id, void **req)
{
    H5VL_log_t *type = (H5VL_log_t *)dt;

    assert(type->under_object);

    H5VLdatatype_close(type->under_object, native_plugin_id, dxpl_id, req);
    free(type);

    printf("------- LOG H5Tclose\n");
    return 1;
}

static void *
H5VL_log_object_open(void *obj, H5VL_loc_params_t loc_params, H5I_type_t *opened_type, hid_t dxpl_id, void **req)
{
    H5VL_log_t *new_obj;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    new_obj = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    new_obj->under_object = H5VLobject_open(o->under_object, loc_params, native_plugin_id, opened_type, dxpl_id, req);

    printf("------- LOG H5Oopen\n");
    return (void *)new_obj;
}

static herr_t 
H5VL_log_object_specific(void *obj, H5VL_loc_params_t loc_params, H5VL_object_specific_t specific_type, 
                         hid_t dxpl_id, void **req, va_list arguments)
{
    H5VL_log_t *o = (H5VL_log_t *)obj;

    H5VLobject_specific(o->under_object, loc_params, native_plugin_id, specific_type, dxpl_id, req, arguments);

    printf("------- LOG Object specific\n");
    return 1;
}

static void *
H5VL_log_dataset_create(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req) 
{
    H5VL_log_t *dset;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dset = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dset->under_object = H5VLdataset_create(o->under_object, loc_params, native_plugin_id, name, dcpl_id,  dapl_id, dxpl_id, req);

    printf("------- LOG H5Dcreate\n");
    return (void *)dset;
}

static void *
H5VL_log_dataset_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *dset;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dset = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dset->under_object = H5VLdataset_open(o->under_object, loc_params, native_plugin_id, name, dapl_id, dxpl_id, req);

    printf("------- LOG H5Dopen\n");
    return (void *)dset;
}

static herr_t 
H5VL_log_dataset_read(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                      hid_t file_space_id, hid_t plist_id, void *buf, void **req)
{
    H5VL_log_t *d = (H5VL_log_t *)dset;

    H5VLdataset_read(d->under_object, native_plugin_id, mem_type_id, mem_space_id, file_space_id, 
                     plist_id, buf, req);

    printf("------- LOG H5Dread\n");
    return 1;
}
static herr_t 
H5VL_log_dataset_write(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                       hid_t file_space_id, hid_t plist_id, const void *buf, void **req)
{
    H5VL_log_t *d = (H5VL_log_t *)dset;

    H5VLdataset_write(d->under_object, native_plugin_id, mem_type_id, mem_space_id, file_space_id, 
                     plist_id, buf, req);

    printf("------- LOG H5Dwrite\n");
    return 1;
}
static herr_t 
H5VL_log_dataset_close(void *dset, hid_t dxpl_id, void **req)
{
    H5VL_log_t *d = (H5VL_log_t *)dset;

    H5VLdataset_close(d->under_object, native_plugin_id, dxpl_id, req);
    free(d);

    printf("------- LOG H5Dclose\n");
    return 1;
}
\end{lstlisting}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: