### 2015-05-21 WhitePapers/revise_chunks/pics/scaled.tex

\tikzstyle{trans} = [>=stealth, thick, text centered, font=\small]
\tikzstyle{computBox} = [draw=black, thick, inner sep=4pt, inner ysep=4pt,
  rectangle, rounded corners]

\newcommand{\drawcuboid}[5]{% width, height, depth, scale, color
\begin{scope}[scale=#4]
\coordinate (O) at (0,0,0);
\coordinate (A) at (0,#2,0);
\coordinate (B) at (0,#2,#3);
\coordinate (C) at (0,0,#3);
\coordinate (D) at (#1,0,0);
\coordinate (E) at (#1,#2,0);
\coordinate (F) at (#1,#2,#3);
\coordinate (G) at (#1,0,#3);

\draw[fill=#5] (O) -- (C) -- (G) -- (D) -- cycle;% Bottom Face
\draw[fill=#5] (O) -- (A) -- (E) -- (D) -- cycle;% Back Face
\draw[fill=#5] (O) -- (A) -- (B) -- (C) -- cycle;% Left Face
\draw[fill=#5] (D) -- (E) -- (F) -- (G) -- cycle;% Right Face
\draw[fill=#5] (C) -- (B) -- (F) -- (G) -- cycle;% Front Face
\draw[fill=#5] (A) -- (B) -- (F) -- (E) -- cycle;% Top Face

\draw[dashed] (O) -- (D);
\draw[dashed] (O) -- (C);
\draw[dashed] (O) -- (A);

\coordinate[left=6pt of A] (A1);
\coordinate[left=6pt of B] (B1);
\coordinate[left=6pt of C] (C1);
\coordinate[below=4pt of C] (C2);
\coordinate[below=4pt of G] (G1);
\draw[<->,trans] (C2) -- node [below] {#1} (G1);
\draw[<->,trans] (C1) -- node [left] {#2} (B1);
\draw[<->,trans] (A1) -- node [above left] {#3} (B1);

\end{scope}
}

%\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

\begin{tikzpicture}

\node (cubes) {
\begin{minipage}{0.32\textwidth}
\begin{tikzpicture}
% Origin axes
\draw[->,draw=paperblue,thick] (0,0,0) -- (3.2,0,0);
\draw[->,draw=paperblue,thick] (0,0,0) -- (0,3.8,0);
\draw[->,draw=paperblue,thick] (0,0,0) -- (0,0,4);
\node[font=\itshape] at (3.4,0,0) {x};
\node[font=\itshape] at (0,4,0) {y};
\node[font=\itshape] at (0,0,4.2) {z};

% Cubes
\drawcuboid{60}{75}{75}{0.04}{white}
\drawcuboid{10}{25}{25}{0.04}{paperblue!20}
\end{tikzpicture}
\end{minipage}
};

% Hash value computation
\node [computBox, font=\footnotesize, text width=100pt, right=50pt of cubes] (scaled) {
  \begin{minipage}{\textwidth}
  {
\textbf{Scaled coordinates:}\\
\begin{math}
   X/x = 60/10 = 6\\
   Y/y = 75/25 = 3\\
   Z/z = 75/25 = 3
\end{math}

\smallskip
\textbf{Number of bits to represent\\ coordinates:}\\
\begin{math}
   \lceil \log_2(X/x) \rceil = 3\\
   \lceil \log_2(Y/y) \rceil = 2\\
   \lceil \log_2(Z/z) \rceil = 2
\end{math}
  }
  \end{minipage}
};

\node [right=of scaled,rectangle split,rectangle split horizontal,
  rectangle split parts=2,draw, minimum height=15pt,text centered,anchor=center,font=\small] (zhash) {0 \nodepart{two} 0};
\node [right=14pt of zhash,rectangle split,rectangle split horizontal,
  rectangle split parts=2,draw, minimum height=15pt,text centered,anchor=center,font=\small] (yhash) {0 \nodepart{two} 0};
\node [right=20pt of yhash,rectangle split,rectangle split horizontal,
  rectangle split parts=3,draw, minimum height=15pt,text centered,anchor=center,font=\small] (xhash) {0 \nodepart{two} 0 \nodepart{three} 0};
\node [below=2pt of xhash] {$x$};
\node [below=2pt of yhash] {$y$};
\node [below=2pt of zhash] {$z$};

\draw[very thick, ->] (cubes) -- node[above, text width=4cm, align=center, font=\bfseries]
    {Hash-value} ($ (scaled.west) - (0.5,0) $);

\end{tikzpicture}