### 2015-07-20 WhitePapers/revise_chunks_seminar/pics/chunked.tex

\input{pics/common}

\begin{tikzpicture}

\node[inner sep=0] (dset) {};
\node[text height=25pt, left=23pt of dset, inner sep=0] (chunk0) {};
\drawdataset{shift=(chunk0.south), scale=0.5}{3}{0}{0}
\node[text height=25pt, right=20pt of dset, inner sep=0] (chunk1) {};
\drawdataset{shift=(chunk1.south), scale=0.5}{3}{-1}{3}
\node[text height=25pt, below=20pt of chunk0, inner sep=0] (chunk2) {};
\drawdataset{shift=(chunk2.south), scale=0.5}{3}{3}{-1}
\node[text height=25pt, below=20pt of chunk1, inner sep=0] (chunk3) {};
\drawdataset{shift=(chunk3.south), scale=0.5}{3}{-1}{-1}

\end{tikzpicture}