### 2015-07-20 WhitePapers/revise_chunks_seminar/slides.tex

\documentclass[10pt,first,firstsupp,handout]{THG_slides}

% Options for beamer:
%
% 9,10,11,12,13,14,17pt  Fontsizes
% 
% compress: navigation bar becomes smaller
% t       : place contents of frames on top (alternative: b,c)
% handout : handoutversion
% notes   : show notes
% notes=onlyslideswithnotes
%

\setbeamertemplate{note page}{\ \\[.3cm]
\textbf{\color{blue}Notes:}\\%[0.1cm]
{\footnotesize %\tiny
\insertnote}}

% URLs
\usepackage{url}

%% Spacing adjustments
\usepackage{textcomp} % Must be loaded to prevent conflict microtype/siunitx
\usepackage{microtype}

% Nicer tabs
\usepackage{booktabs}

%% TikZ
\usepackage{tikz}
\usepackage{gnuplot-lua-tikz}
\usetikzlibrary{tikzmark, calc, patterns, shapes, positioning,arrows,
  decorations.pathmorphing, decorations.pathreplacing, decorations.text,
  math, trees, shadows}

% Tikz macros
\tikzset{%
  highlight/.style={rectangle,rounded corners,draw,thick}
}
\newcommand{\tablearrowright}[4]{%
  \tikz[overlay,remember picture]{
  \draw ($(pic cs:#1) + (5pt,4pt)$) edge[bend right=45,-stealth]
        ($(pic cs:#2) + (5pt,4pt)$)
        node [highlight,right,xshift=5pt,yshift=7pt,#4] {#3};}
}
\newcommand{\tablearrowleft}[4]{%
  \tikz[overlay,remember picture]{
  \draw ($(pic cs:#1) + (5pt,2pt)$) edge[bend left=45,-stealth]
        ($(pic cs:#2) + (5pt,2pt)$)
        node [highlight,right,xshift=5pt,yshift=-7pt,#4] {#3};}
}
\newcommand*{\yellowemph}[1]{%
  \tikz[baseline=(X.base)] \node[rectangle, fill=yellow, rounded corners, inner sep=0.3mm] (X) {#1};%
}
\newcommand{\smiley}{%
  \tikz[baseline=-0.75ex,black]{
    \draw circle (2mm);
    \node[fill,circle,inner sep=0.5pt] (left eye) at (135:0.8mm) {};
    \node[fill,circle,inner sep=0.5pt] (right eye) at (45:0.8mm) {};
    \draw (-145:0.9mm) arc (-120:-60:1.5mm);
  }
}

% Additional colors
\usepackage{xcolor}
\colorlet{paperblue}{-red!75!green!50}

% SI units
\usepackage{siunitx}
\sisetup{per-mode = symbol, binary-units = true}
\providecommand{\gbps}[1]{\SI{#1}{\giga\byte\per\second}}
\providecommand{\mbps}[1]{\SI{#1}{\mega\byte\per\second}}
\providecommand{\gb}[1]{\SI{#1}{\giga\byte}}
\providecommand{\mb}[1]{\SI{#1}{\mega\byte}}
\providecommand{\kb}[1]{\SI{#1}{\kilo\byte}}

%\usepackage[format=hang,labelfont=bf,labelsep=space,font=small]{caption}

% Code snippets
\usepackage{listings}
\def\lstsetc{\lstset{language=C,
%  numbers=left,
%  xleftmargin=2.2em,
  xleftmargin=1.8em,
  framexleftmargin=0.2em,
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  showspaces=false, 
  showstringspaces=false,
  breaklines=true,
  basicstyle=\footnotesize\ttfamily,
  stringstyle=\itshape,
  commentstyle=\itshape\bfseries,
  frame=leftline,
  morekeywords={main, size_t, malloc, free, write, H5Screate_simple},
  morekeywords={[2]H5S_UNLIMITED},
  keywordstyle={[2]\color{gray}\bfseries}
  }
}

% Shows only the navigational symbol for navigating frames.
\setbeamertemplate{navigation symbols}[only frame symbol]

\setlayoutscale{0.5}
\setparametertextfont{\scriptsize}
\setlabelfont{\scriptsize}

\AtBeginSection[]
{
  \frame{
    \frametitle{Outline}
    \tableofcontents[currentsection]
  }
}

\begin{document}

\title{Extending HDF5 Datasets: Enhancements to the Chunk Indexing Methods}
\author{Vailin Choi, Jerome Soumagne, Quincey Koziol}
\institute{The HDF Group}

\begin{frame}
\maketitle
\end{frame}

\input{slides-body}

\end{document}