### 2015-10-15 RFCs/HDF5_Library/AnalysisExt/pics/hdf5_view.tex

\tikzstyle{groupBox} = [draw, ellipse, fill=white, minimum size=20pt,
  text centered, font=\footnotesize\sffamily, text width=20pt, inner sep=0]

\tikzstyle{textBox} = [fill=white, inner sep=2pt, text centered, font=\footnotesize\sffamily]

\tikzstyle{link} = [->, thick, >=stealth', shorten <=1pt, shorten >=1pt]

\pgfmathsetseed{1}

\def\drawdataset#1{%
\begin{scope}[#1]
  \foreach \y in {-0.5} {
      \foreach \x in {-2.5,-1.5,-0.5,0.5,1.5} {
          \path[fill=white, draw] (\x,\y) rectangle ++(1,1);
      }
  }
\end{scope}
}

\begin{tikzpicture}

% View
\node[groupBox] (view) {/};

% Refs
\node[below=30pt of view, inner sep=0] (viewgroup) {};

\node[rectangle, text height=7pt, left=70pt of viewgroup, inner sep=0] (obj_ref) {};
\drawdataset{shift=(obj_ref.south), scale=0.3}
\node[rectangle, text height=7pt, below=10pt of viewgroup, inner sep=0] (reg_ref) {};
\drawdataset{shift=(reg_ref), scale=0.3}
\node[rectangle, text height=7pt, right=70pt of viewgroup, inner sep=0] (attr_ref) {};
\drawdataset{shift=(attr_ref.south), scale=0.3}
\node[left=12pt of obj_ref] (g1) {};
\node[right=12pt of attr_ref] (g2) {};

% Links
\draw[link] (view) -- node [textBox] {Obj\_refs} (obj_ref);
\draw[link] (view) -- node [textBox] {Reg\_refs} (reg_ref);
\draw[link] (view) -- node [textBox] {Attr\_refs} (attr_ref);

\begin{scope}[on background layer, shape aspect=.10]
\tikzset{every node/.style={cylinder, shape border rotate=90, draw}}
\node [fit=(view) (obj_ref) (reg_ref) (attr_ref) (g1) (g2)] {};
\end{scope}

%\node [minimum width=1.5cm] at (1.5,0) {B};
\end{tikzpicture}