### 2015-10-15 RFCs/HDF5_Library/AnalysisExt/analysis_ext.tex
A few fixes to previous commit
Former-commit-id: 66a36b3bc72dbb39f4f492b15a12021995ef594e

\documentclass[letterpaper,hyper]{THG_RFC}

% Path to figures, plots
\graphicspath{{./pics/}{./plots/}}

%% TikZ
\usepackage{tikz}
\usetikzlibrary{patterns, shapes, positioning, decorations.pathreplacing,
backgrounds, fit}
\usepackage{gnuplot-lua-tikz}

% SI units
%\usepackage{textcomp}
\usepackage[binary-units]{siunitx}
\sisetup{per-mode = symbol}
\providecommand{\gbps}[1]{\SI{#1}{\giga\byte\per\second}}
\providecommand{\mbps}[1]{\SI{#1}{\mega\byte\per\second}}
\providecommand{\gb}[1]{\SI{#1}{\giga\byte}}
\providecommand{\mb}[1]{\SI{#1}{\mega\byte}}
\providecommand{\kb}[1]{\SI{#1}{\kilo\byte}}

% Code snippets
\usepackage{listings}
\def\lstsetc{\lstset{language=C,
  numbers=left,
  xleftmargin=1.8em,
  frame=L,
  framexleftmargin=0.2em,
  rulecolor=\color{rulecolor},
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  showspaces=false, 
  showstringspaces=false,
  breaklines=true,
  basicstyle=\footnotesize\ttfamily,
  stringstyle=\itshape,
  commentstyle=\itshape\bfseries,
  morekeywords={main, size_t, malloc, free,
    hsize_t, hid_t, herr_t,
    H5X_class_t, H5X_type_t, H5X_info_t,
    H5Q_combine_op_t, H5Q_type_t, H5Q_match_op_t,
    H5Fcreate, H5Fclose,
    H5Dcreate, H5Dopen, H5Dclose, H5Dwrite, H5Dread, H5Dquery,
    H5Qcreate, H5Qclose, H5Qcombine, H5Qget_type, H5Qget_match_op,
    H5Qget_components, H5Qget_combine_op, H5Qencode, H5Qdecode,
    H5Qapply,
    H5Screate_simple, H5Sclose, H5Sget_select_npoints,
    H5Xregister, H5Xunregister, H5Xcreate, H5Xremove, H5Xget_count, H5Xget_size,
    H5Xget_info
    }
  }
}

% Title, author, etc
\title{Data Analysis Extensions}
\author{Jerome Soumagne}
\author{Quincey Koziol}
\date{Oct 14, 2015}
\rfcversion{2014-07-17.v6}
\revision{Jul. 17, 2014}{Version 1 circulated for comment within The HDF Group.}
\revision{Aug. 20, 2014}{Version 2 circulated for comment within The HDF Group.}
\revision{Nov. 14, 2014}{Version 3 circulated for comment within The HDF Group.}
\revision{Feb. 25, 2015}{Version 4 edits and typo fixing.}
\revision{Jul. 24, 2015}{Version 5 edits to view object.}
\revision{Oct. 14, 2015}{Version 6 rework of query/view sections.}

%% Start the document
\begin{document}

%% Title
\maketitle

%% Abstract
\begin{abstract}
Accessing, selecting and retrieving data from an HDF5 container can be a time
consuming process, particularly so when data is very large. To enable, ease and 
accelerate this process, we introduce in this RFC extensions to the library
to efficiently query, select and index data.
\end{abstract}

\input{analysis_ext-body}

\section*{Revision History}
\makerevisions

%% References
\bibliographystyle{abbrv}
\bibliography{analysis_ext}

\end{document}