### 2018-01-11 RFCs/HDF5/VOL/user_guide/appendix.tex
misc. doc changes
Former-commit-id: 06e7d1ed7bd432a0564b1d363fe2435226459869

\begin{appendices}
\section{\\VOL Plugin Example}
\label{sec:A}

This appendix shows a simple implementation of a {\tt printf} plugin that stacks on top of the native HDF5 plugin. Note that not all callbacks have been implemented since this is just an example for explaining the general framework of the VOL plugin.

\begin{lstlisting}

#include <stdio.h>
#include <stdlib.h>
#include <assert.h>
#include "hdf5.h"

#define LOG 502

static herr_t H5VL_log_init(hid_t vipl_id);
static herr_t H5VL_log_term(hid_t vtpl_id);

/* Datatype callbacks */
static void *H5VL_log_datatype_commit(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id, hid_t dxpl_id, void **req);
static void *H5VL_log_datatype_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t tapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_datatype_get(void *dt, H5VL_datatype_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
static herr_t H5VL_log_datatype_close(void *dt, hid_t dxpl_id, void **req);

/* Dataset callbacks */
static void *H5VL_log_dataset_create(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req);
static void *H5VL_log_dataset_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_dataset_read(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                                    hid_t file_space_id, hid_t plist_id, void *buf, void **req);
static herr_t H5VL_log_dataset_write(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                                     hid_t file_space_id, hid_t plist_id, const void *buf, void **req);
static herr_t H5VL_log_dataset_close(void *dset, hid_t dxpl_id, void **req);

/* File callbacks */
static void *H5VL_log_file_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id, void **req);
static void *H5VL_log_file_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_file_get(void *file, H5VL_file_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
static herr_t H5VL_log_file_close(void *file, hid_t dxpl_id, void **req);

/* Group callbacks */
static void *H5VL_log_group_create(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id, void **req);
static herr_t H5VL_log_group_close(void *grp, hid_t dxpl_id, void **req);

/* Link callbacks */

/* Object callbacks */
static void *H5VL_log_object_open(void *obj, H5VL_loc_params_t loc_params, H5I_type_t *opened_type, hid_t dxpl_id, void **req);
static herr_t H5VL_log_object_specific(void *obj, H5VL_loc_params_t loc_params, H5VL_object_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);

hid_t native_plugin_id = -1;

static const H5VL_class_t H5VL_log_g = {
    0,
    LOG,
    "log",					/* name */
    H5VL_log_init,                              /* initialize */
    H5VL_log_term,                              /* terminate */
    sizeof(hid_t),
    NULL,
    NULL,
    {                                           /* attribute_cls */
        NULL, //H5VL_log_attr_create,                /* create */
        NULL, //H5VL_log_attr_open,                  /* open */
        NULL, //H5VL_log_attr_read,                  /* read */
        NULL, //H5VL_log_attr_write,                 /* write */
        NULL, //H5VL_log_attr_get,                   /* get */
        NULL, //H5VL_log_attr_specific,              /* specific */
        NULL, //H5VL_log_attr_optional,              /* optional */
        NULL  //H5VL_log_attr_close                  /* close */
    },
    {                                           /* dataset_cls */
        H5VL_log_dataset_create,                    /* create */
        H5VL_log_dataset_open,                      /* open */
        H5VL_log_dataset_read,                      /* read */
        H5VL_log_dataset_write,                     /* write */
        NULL, //H5VL_log_dataset_get,               /* get */
        NULL, //H5VL_log_dataset_specific,          /* specific */
        NULL, //H5VL_log_dataset_optional,          /* optional */
        H5VL_log_dataset_close                      /* close */
    },
    {                                               /* datatype_cls */
        H5VL_log_datatype_commit,                   /* commit */
        H5VL_log_datatype_open,                     /* open */
        H5VL_log_datatype_get,                      /* get_size */
        NULL, //H5VL_log_datatype_specific,         /* specific */
        NULL, //H5VL_log_datatype_optional,         /* optional */
        H5VL_log_datatype_close                     /* close */
    },
    {                                           /* file_cls */
        H5VL_log_file_create,                      /* create */
        H5VL_log_file_open,                        /* open */
        H5VL_log_file_get,                         /* get */
        NULL, //H5VL_log_file_specific,            /* specific */
        NULL, //H5VL_log_file_optional,            /* optional */
        H5VL_log_file_close                        /* close */
    },
    {                                           /* group_cls */
        H5VL_log_group_create,                     /* create */
        NULL, //H5VL_log_group_open,               /* open */
        NULL, //H5VL_log_group_get,                /* get */
        NULL, //H5VL_log_group_specific,           /* specific */
        NULL, //H5VL_log_group_optional,           /* optional */
        H5VL_log_group_close                       /* close */
    },
    {                                           /* link_cls */
        NULL, //H5VL_log_link_create,                /* create */
        NULL, //H5VL_log_link_copy,                  /* copy */
        NULL, //H5VL_log_link_move,                  /* move */
        NULL, //H5VL_log_link_get,                   /* get */
        NULL, //H5VL_log_link_specific,              /* specific */
        NULL, //H5VL_log_link_optional,              /* optional */
    },
    {                                           /* object_cls */
        H5VL_log_object_open,                        /* open */
        NULL, //H5VL_log_object_copy,                /* copy */
        NULL, //H5VL_log_object_get,                 /* get */
        H5VL_log_object_specific,                    /* specific */
        NULL, //H5VL_log_object_optional,            /* optional */
    },
    {
        NULL,
        NULL,
        NULL
    },
    NULL
};

typedef struct H5VL_log_t {
    void   *under_object;
} H5VL_log_t;

static herr_t
visit_cb(hid_t oid, const char *name,
         const H5O_info_t *oinfo, void *udata)
{
    ssize_t len;
    char n[25];

    if(H5Iget_type(oid) == H5I_GROUP) {
        len = H5VLget_plugin_name(oid, n, 50);
        printf ("Visiting GROUP VOL name = %s  %d\n", n, len);
    }
    if(H5Iget_type(oid) == H5I_DATASET) 
        printf("visiting dataset\n");
    if(H5Iget_type(oid) == H5I_DATATYPE) 
        printf("visiting datatype\n");

    return 1;
} /* end h5_verify_cached_stabs_cb() */

int main(int argc, char **argv) {
        const char file_name[]="large_dataset.h5";
	const char group_name[]="/Group";
	const char dataset_name[]="Data";
	char fullpath[500];
	hid_t file_id;
	hid_t group_id;
	hid_t dataspaceId;
        hid_t datasetId;
        hid_t acc_tpl;
        hid_t under_fapl;
        hid_t vol_id, vol_id2;
        hid_t int_id;
        hid_t attr;
        hid_t space;
	const unsigned int nelem=60;
	int *data;
	unsigned int i;
	hsize_t dims[1];
        ssize_t len;
        char name[25];
        static hsize_t      ds_size[2] = {10, 20};

        under_fapl = H5Pcreate (H5P_FILE_ACCESS);
        H5Pset_fapl_native(under_fapl);
        assert(H5VLis_registered("native") == 1);

        vol_id = H5VLregister (&H5VL_log_g);
        assert(vol_id > 0);
        assert(H5VLis_registered("log") == 1);

        vol_id2 = H5VLget_plugin_id("log");
        H5VLinitialize(vol_id2, H5P_DEFAULT);
        H5VLclose(vol_id2);

        native_plugin_id = H5VLget_plugin_id("native");
        assert(native_plugin_id > 0);

        acc_tpl = H5Pcreate (H5P_FILE_ACCESS);
        H5Pset_vol(acc_tpl, vol_id, &under_fapl);

	file_id = H5Fcreate(file_name, H5F_ACC_TRUNC, H5P_DEFAULT, acc_tpl);
        len = H5VLget_plugin_name(file_id, name, 25);
        printf ("FILE VOL name = %s  %d\n", name, len);

	group_id = H5Gcreate2(file_id, group_name, H5P_DEFAULT, H5P_DEFAULT, H5P_DEFAULT);
        len = H5VLget_plugin_name(group_id, name, 50);
        printf ("GROUP VOL name = %s  %d\n", name, len);

        int_id = H5Tcopy(H5T_NATIVE_INT);
        H5Tcommit2(file_id, "int", int_id, H5P_DEFAULT, H5P_DEFAULT, H5P_DEFAULT);
        len = H5VLget_plugin_name(int_id, name, 50);
        printf ("DT COMMIT name = %s  %d\n", name, len);
        H5Tclose(int_id);

        int_id = H5Topen2(file_id, "int", H5P_DEFAULT);
        len = H5VLget_plugin_name(int_id, name, 50);
        printf ("DT OPEN name = %s  %d\n", name, len);
        H5Tclose(int_id);

        int_id = H5Oopen(file_id,"int",H5P_DEFAULT);
        len = H5VLget_plugin_name(int_id, name, 50);
        printf ("DT OOPEN name = %s  %d\n", name, len);

        len = H5Fget_name(file_id, name, 50);
        printf("name = %d  %s\n", len, name);

	data = malloc (sizeof(int)*nelem);
	for(i=0;i<nelem;++i)
	  data[i]=i;

	dims [0] = 60;
	dataspaceId = H5Screate_simple(1, dims, NULL); 
        space = H5Screate_simple (2, ds_size, ds_size);

	sprintf(fullpath,"%s/%s",group_name,dataset_name);
	datasetId = H5Dcreate2(file_id, fullpath, H5T_NATIVE_INT, dataspaceId, H5P_DEFAULT, H5P_DEFAULT, H5P_DEFAULT);
	H5Sclose(dataspaceId);

        len = H5VLget_plugin_name(datasetId, name, 50);
        printf ("DSET name = %s  %d\n", name, len);

	H5Dwrite(datasetId, H5T_NATIVE_INT, H5S_ALL, H5S_ALL, H5P_DEFAULT, data);
	H5Dclose(datasetId);

        H5Ovisit(file_id, H5_INDEX_NAME, H5_ITER_NATIVE, visit_cb, NULL);

	free (data);
        H5Oclose(int_id);
        H5Sclose (space);
	H5Gclose(group_id);

	H5Fclose(file_id);
        H5Pclose(acc_tpl);
        H5Pclose(under_fapl);

        H5VLclose(native_plugin_id);
        H5VLterminate(vol_id, H5P_DEFAULT);
        H5VLunregister (vol_id);
        assert(H5VLis_registered("log") == 0);
	return 0;
}

static herr_t H5VL_log_init(hid_t vipl_id)
{
    printf("------- LOG INIT\n");
    return 0;
}

static herr_t H5VL_log_term(hid_t vtpl_id)
{
    printf("------- LOG TERM\n");
    return 0;
}

static void *
H5VL_log_file_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id, void **req)
{
    hid_t under_fapl;
    H5VL_log_t *file;

    file = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    under_fapl = *((hid_t *)H5Pget_vol_info(fapl_id));
    file->under_object = H5VLfile_create(name, flags, fcpl_id, under_fapl, dxpl_id, req);

    printf("------- LOG H5Fcreate\n");
    return (void *)file;
}

static void *
H5VL_log_file_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req)
{
    hid_t under_fapl;
    H5VL_log_t *file;

    file = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    under_fapl = *((hid_t *)H5Pget_vol_info(fapl_id));
    file->under_object = H5VLfile_open(name, flags, under_fapl, dxpl_id, req);

    printf("------- LOG H5Fopen\n");
    return (void *)file;
}

static herr_t 
H5VL_log_file_get(void *file, H5VL_file_get_t get_type, hid_t dxpl_id, void **req, va_list arguments)
{
    H5VL_log_t *f = (H5VL_log_t *)file;

    H5VLfile_get(f->under_object, native_plugin_id, get_type, dxpl_id, req, arguments);

    printf("------- LOG H5Fget %d\n", get_type);
    return 1;
}
static herr_t 
H5VL_log_file_close(void *file, hid_t dxpl_id, void **req)
{
    H5VL_log_t *f = (H5VL_log_t *)file;

    H5VLfile_close(f->under_object, native_plugin_id, dxpl_id, req);
    free(f);

    printf("------- LOG H5Fclose\n");
    return 1;
}

static void *
H5VL_log_group_create(void *obj, H5VL_loc_params_t loc_params, const char *name, 
                      hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *group;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    group = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    group->under_object = H5VLgroup_create(o->under_object, loc_params, native_plugin_id, name, gcpl_id,  gapl_id, dxpl_id, req);

    printf("------- LOG H5Gcreate\n");
    return (void *)group;
}

static herr_t 
H5VL_log_group_close(void *grp, hid_t dxpl_id, void **req)
{
    H5VL_log_t *g = (H5VL_log_t *)grp;

    H5VLgroup_close(g->under_object, native_plugin_id, dxpl_id, req);
    free(g);

    printf("------- LOG H5Gclose\n");
    return 1;
}

static void *
H5VL_log_datatype_commit(void *obj, H5VL_loc_params_t loc_params, const char *name, 
                         hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *dt;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dt = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dt->under_object = H5VLdatatype_commit(o->under_object, loc_params, native_plugin_id, name, 
                                           type_id, lcpl_id, tcpl_id, tapl_id, dxpl_id, req);

    printf("------- LOG H5Tcommit\n");
    return dt;
}
static void *
H5VL_log_datatype_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t tapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *dt;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dt = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dt->under_object = H5VLdatatype_open(o->under_object, loc_params, native_plugin_id, name, tapl_id, dxpl_id, req);

    printf("------- LOG H5Topen\n");
    return (void *)dt;
}

static herr_t 
H5VL_log_datatype_get(void *dt, H5VL_datatype_get_t get_type, hid_t dxpl_id, void **req, va_list arguments)
{
    H5VL_log_t *o = (H5VL_log_t *)dt;
    herr_t ret_value;

    ret_value = H5VLdatatype_get(o->under_object, native_plugin_id, get_type, dxpl_id, req, arguments);

    printf("------- LOG datatype get\n");
    return ret_value;
}

static herr_t 
H5VL_log_datatype_close(void *dt, hid_t dxpl_id, void **req)
{
    H5VL_log_t *type = (H5VL_log_t *)dt;

    assert(type->under_object);

    H5VLdatatype_close(type->under_object, native_plugin_id, dxpl_id, req);
    free(type);

    printf("------- LOG H5Tclose\n");
    return 1;
}

static void *
H5VL_log_object_open(void *obj, H5VL_loc_params_t loc_params, H5I_type_t *opened_type, hid_t dxpl_id, void **req)
{
    H5VL_log_t *new_obj;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    new_obj = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    new_obj->under_object = H5VLobject_open(o->under_object, loc_params, native_plugin_id, opened_type, dxpl_id, req);

    printf("------- LOG H5Oopen\n");
    return (void *)new_obj;
}

static herr_t 
H5VL_log_object_specific(void *obj, H5VL_loc_params_t loc_params, H5VL_object_specific_t specific_type, 
                         hid_t dxpl_id, void **req, va_list arguments)
{
    H5VL_log_t *o = (H5VL_log_t *)obj;

    H5VLobject_specific(o->under_object, loc_params, native_plugin_id, specific_type, dxpl_id, req, arguments);

    printf("------- LOG Object specific\n");
    return 1;
}

static void *
H5VL_log_dataset_create(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req) 
{
    H5VL_log_t *dset;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dset = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dset->under_object = H5VLdataset_create(o->under_object, loc_params, native_plugin_id, name, dcpl_id,  dapl_id, dxpl_id, req);

    printf("------- LOG H5Dcreate\n");
    return (void *)dset;
}

static void *
H5VL_log_dataset_open(void *obj, H5VL_loc_params_t loc_params, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req)
{
    H5VL_log_t *dset;
    H5VL_log_t *o = (H5VL_log_t *)obj;

    dset = (H5VL_log_t *)calloc(1, sizeof(H5VL_log_t));

    dset->under_object = H5VLdataset_open(o->under_object, loc_params, native_plugin_id, name, dapl_id, dxpl_id, req);

    printf("------- LOG H5Dopen\n");
    return (void *)dset;
}

static herr_t 
H5VL_log_dataset_read(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                      hid_t file_space_id, hid_t plist_id, void *buf, void **req)
{
    H5VL_log_t *d = (H5VL_log_t *)dset;

    H5VLdataset_read(d->under_object, native_plugin_id, mem_type_id, mem_space_id, file_space_id, 
                     plist_id, buf, req);

    printf("------- LOG H5Dread\n");
    return 1;
}
static herr_t 
H5VL_log_dataset_write(void *dset, hid_t mem_type_id, hid_t mem_space_id,
                       hid_t file_space_id, hid_t plist_id, const void *buf, void **req)
{
    H5VL_log_t *d = (H5VL_log_t *)dset;

    H5VLdataset_write(d->under_object, native_plugin_id, mem_type_id, mem_space_id, file_space_id, 
                     plist_id, buf, req);

    printf("------- LOG H5Dwrite\n");
    return 1;
}
static herr_t 
H5VL_log_dataset_close(void *dset, hid_t dxpl_id, void **req)
{
    H5VL_log_t *d = (H5VL_log_t *)dset;

    H5VLdataset_close(d->under_object, native_plugin_id, dxpl_id, req);
    free(d);

    printf("------- LOG H5Dclose\n");
    return 1;
}
\end{lstlisting}

\section{\\Listing of Direct Mapping to VOL Callback APIs}
\label{sec:B}

\begin{lstlisting}
/* ATTRIBUTE OBJECT ROUTINES */
void *H5VLattr_create(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *attr_name, hid_t acpl_id, hid_t aapl_id, hid_t dxpl_id, void **req);
void *H5VLattr_open(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t aapl_id, hid_t dxpl_id, void **req);
herr_t H5VLattr_read(void *attr, hid_t plugin_id, hid_t dtype_id, void *buf, hid_t dxpl_id, void **req);
herr_t H5VLattr_write(void *attr, hid_t plugin_id, hid_t dtype_id, const void *buf, hid_t dxpl_id, void **req);
herr_t H5VLattr_get(void *obj, hid_t plugin_id, H5VL_attr_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLattr_specific(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, H5VL_attr_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLattr_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLattr_close(void *attr, hid_t plugin_id, hid_t dxpl_id, void **req);

/* DATASET OBJECT ROUTINES */
void *H5VLdataset_create(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req);
void *H5VLdataset_open(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
herr_t H5VLdataset_read(void *dset, hid_t plugin_id, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t plist_id, void *buf, void **req);
herr_t H5VLdataset_write(void *dset, hid_t plugin_id, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t plist_id, const void *buf, void **req);
herr_t H5VLdataset_get(void *dset, hid_t plugin_id, H5VL_dataset_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdataset_specific(void *obj, hid_t plugin_id, H5VL_dataset_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdataset_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdataset_close(void *dset, hid_t plugin_id, hid_t dxpl_id, void **req);

/* DATATYPE OBJECT ROUTINES */
void *H5VLdatatype_commit(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id, hid_t dxpl_id, void **req);
void *H5VLdatatype_open(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t tapl_id, hid_t dxpl_id, void **req);
herr_t H5VLdatatype_get(void *dt, hid_t plugin_id, H5VL_datatype_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdatatype_specific(void *obj, hid_t plugin_id, H5VL_datatype_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdatatype_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdatatype_close(void *dt, hid_t plugin_id, hid_t dxpl_id, void **req);

/* FILE OBJECT ROUTINES */
void *H5VLfile_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id, void **req);
void *H5VLfile_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req);
herr_t H5VLfile_get(void *file, hid_t plugin_id, H5VL_file_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLfile_specific(void *obj, hid_t plugin_id, H5VL_file_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLfile_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLfile_close(void *file, hid_t plugin_id, hid_t dxpl_id, void **req);

/* GROUP OBJECT ROUTINES */
void *H5VLgroup_create(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id, void **req);
void *H5VLgroup_open(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, const char *name, hid_t gapl_id, hid_t dxpl_id, void **req);
herr_t H5VLgroup_get(void *obj, hid_t plugin_id, H5VL_group_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLgroup_specific(void *obj, hid_t plugin_id, H5VL_group_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLgroup_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLgroup_close(void *grp, hid_t plugin_id, hid_t dxpl_id, void **req);

/* LINK OBJECT ROUTINES */
herr_t H5VLlink_create(H5VL_link_create_type_t create_type, void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_copy(void *src_obj, H5VL_loc_params_t loc_params1, void *dst_obj, H5VL_loc_params_t loc_params2, hid_t plugin_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_move(void *src_obj, H5VL_loc_params_t loc_params1, void *dst_obj, H5VL_loc_params_t loc_params2, hid_t plugin_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_get(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, H5VL_link_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLlink_specific(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, H5VL_link_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLlink_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);

/* OBJECT ROUTINES */
void *H5VLobject_open(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, H5I_type_t *opened_type, hid_t dxpl_id, void **req);
herr_t H5VLobject_copy(void *src_obj, H5VL_loc_params_t loc_params1, hid_t plugin_id1, const char *src_name,                      void *dst_obj, H5VL_loc_params_t loc_params2, hid_t plugin_id2, const char *dst_name, hid_t ocpypl_id, hid_t lcpl_id, hid_t dxpl_id, void **req);
herr_t H5VLobject_get(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, H5VL_object_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLobject_specific(void *obj, H5VL_loc_params_t loc_params, hid_t plugin_id, H5VL_object_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLobject_optional(void *obj, hid_t plugin_id, hid_t dxpl_id, void **req, va_list arguments);

/* ASYNCHRONOUS ROUTINES */
herr_t H5VLrequest_cancel(void **req, hid_t plugin_id, H5ES_status_t *status);
herr_t H5VLrequest_test(void **req, hid_t plugin_id, H5ES_status_t *status);
herr_t H5VLrequest_wait(void **req, hid_t plugin_id, H5ES_status_t *status);
\end{lstlisting}

\end{appendices}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: