### 2019-10-31 RFCs/HDF5/VOL/user_guide/body.tex
Corrections to some API calls in the big HDF5 API call to VOL callback table. Marked functions as deprected in all tables.
Former-commit-id: 7ab94fde980b16dc90076056810cbaa38213fa85

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

\section{Introduction}

\section{Quickstart}
\subsection*{Use A VOL-Enabled HDF5 Library}
\subsection*{Make Sure The VOL Connector Is In The Search Path}
\subsection*{Update Your Code To Load The Connector}

\section{Loading and Registering Connectors}
\subsection{VOL Connector Search Path}
\subsection{Connector-Specific Registration Calls}
\subsection{\tt H5Pset\_vol()}
\subsection{Environment Variable}

\section{Adapting HDF5 Software to Use the VOL}

\subsection{Specific API Call Substitutions}
${\tt H5Fis\_hdf5} \rightarrow {\tt H5Fis\_accessible}$

${\tt H5Oget\_info(1|2)} \rightarrow {\tt H5Ofuture\_function\_goes\_here}$

\section{Using VOL Connectors With The HDF5 Command-Line Tools}
\subsection{Connection Strings}
\subsection{Connector Compatibility}

\section{Compatibility}
\subsection{Feature Flags}
\begin{tcolorbox}[colback=yellow!10!white,colframe=red!75!black,title=PRELIMINARY]
    This feature is under active development and is either incomplete or may change in behavior before the final release.
\end{tcolorbox}
\subsection{List of HDF5 Native VOL API Calls}

These API calls will probably fail when used with terminal VOL connectors other than the native HDF5 file format connector. Their use should be protected in code that uses arbitrary VOL connectors.
\begin{longtable}{ |>{\raggedright\arraybackslash}p{\linewidth}| }
    \hline
    H5Aget\_num\_attrs (deprecated) \\
    H5Aiterate1 (deprecated) \\
    \hline
    H5Ddebug \\
    H5Dformat\_convert \\
    H5Dget\_chunk\_index\_type \\
    H5Dget\_chunk\_info \\
    H5Dget\_chunk\_info\_by\_coord \\
    H5Dget\_chunk\_storage\_size \\
    H5Dget\_num\_chunks \\
    H5Dread\_chunk H5Dwrite\_chunk \\
    \hline
    H5FD* \\
    \hline
    H5Fclear\_elink\_file\_cache \\
    H5Fformat\_convert \\
    H5Fget\_dset\_no\_attrs\_hint \\
    H5Fget\_eoa H5Fget\_file\_image \\
    H5Fget\_filesize \\
    H5Fget\_free\_sections \\
    H5Fget\_freespace \\
    H5Fget\_info1 (deprecated) \\
    H5Fget\_info2 \\
    H5Fget\_mdc\_config \\
    H5Fget\_mdc\_hit\_rate \\
    H5Fget\_mdc\_image\_info \\
    H5Fget\_mdc\_logging\_status \\
    H5Fget\_mdc\_size \\
    H5Fget\_metadata\_read\_retry\_info \\
    H5Fget\_mpi\_atomicity \\
    H5Fget\_page\_buffering\_stats \\
    H5Fget\_vfd\_handle \\
    H5Fincrement\_filesize \\
    H5Fis\_hdf5 (deprecated) \\
    H5Freset\_mdc\_hit\_rate\_stats \\
    H5Freset\_page\_buffering\_stats \\
    H5Fset\_dset\_no\_attrs\_hint \\
    H5Fset\_latest\_format (deprecated) \\
    H5Fset\_libver\_bounds \\
    H5Fset\_mdc\_config \\
    H5Fset\_mpi\_atomicity \\
    H5Fstart\_mdc\_logging \\
    H5Fstart\_swmr\_write \\
    H5Fstop\_mdc\_logging \\
    \hline
    H5Gget\_comment (deprecated) \\
    H5Giterate (deprecated) \\
    H5Gget\_info \\
    H5Gget\_info\_by\_name \\
    H5Gget\_info\_by\_idx \\
    H5Gget\_objinfo (deprecated) \\
    H5Gget\_objname\_by\_idx (deprecated) \\
    H5Gget\_objtype\_by\_idx (deprecated) \\
    H5Gset\_comment (deprecated) \\
    \hline
    H5Oare\_mdc\_flushes\_disabled \\
    H5Odisable\_mdc\_flushes \\
    H5Oenable\_mdc\_flushes \\
    H5Oget\_comment \\
    H5Oget\_comment\_by\_name \\
    H5Oget\_info\_by\_idx1 (deprecated) \\
    H5Oget\_info\_by\_idx2 \\
    H5Oget\_info\_by\_name1 (deprecated) \\
    H5Oget\_info\_by\_name2 \\
    H5Oget\_info1 (deprecated) \\
    H5Oget\_info2 \\
    H5Oset\_comment \\
    H5Oset\_comment\_by\_name \\
    \hline
\caption{Alphabetical list of HDF5 API calls specific to the native VOL connector}
\end{longtable}

\subsection{List of HDF5 VOL-Independent API Calls}

These HDF5 API calls do not depend on a particular VOL connector being loaded.

\begin{longtable}{ |>{\raggedright\arraybackslash}p{\linewidth}| }
    \hline
    H5* \\
    \hline
    H5Dfill \\
    H5Dgather \\
    H5Diterate \\
    H5Dscatter \\
    H5Dvlen\_reclaim (deprecated) \\
    H5Dvlen\_get\_buf\_size \\
    \hline
    H5E* \\
    H5I* \\
    \hline
    H5Lis\_registered \\
    H5Lregister \\
    H5Lunpack\_elink\_val \\
    H5Lunregister \\
    \hline
    H5PL* \\
    H5P* \\
    H5S* \\
    H5T* (non-committed) \\
    H5VL* \\
    H5Z* \\
    \hline
\caption{Alphabetical list of VOL-independent HDF5 API calls}
\end{longtable}

\subsection{List of HDF5 API Calls By Callback}

\begin{longtable}{ |l|>{\raggedright\arraybackslash}l| }
    \hline
    \multicolumn{1}{|c|}{VOL Callback} & \multicolumn{1}{|c|}{HDF5 API Call} \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}FILE} \\
    \hline
    create & H5Fcreate \\
    \hline
    open & H5Fopen \\
    \hline
    get & H5Fget\_access\_plist \\
        & H5Fget\_create\_plist \\
        & H5Fget\_fileno \\
        & H5Fget\_intent \\
        & H5Fget\_name \\
        & H5Fget\_obj\_count \\
        & H5Fget\_obj\_ids \\
    \hline
    specific & H5Fdelete \\
             & H5Fflush \\
             & H5Fis\_accessible \\
             & H5Fis\_hdf5 (deprecated, hard-coded to use native connector) \\
             & H5Fmount \\
             & H5Freopen \\
             & H5Funmount \\
    \hline
    close & H5Fclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}GROUP} \\
    \hline
    create & H5Gcreate1 (deprecated) \\
           & H5Gcreate2 \\
           & H5Gcreate\_anon \\
    \hline
    open & H5Gopen1 (deprecated) \\
         & H5Gopen2 \\
    \hline
    get & H5Gget\_create\_plist \\
        & H5Gget\_info \\
        & H5Gget\_info\_by\_idx \\
        & H5Gget\_info\_by\_name \\
        & H5Gget\_num\_objs (deprecated) \\
    \hline
    specific & H5Gflush \\
             & H5Grefresh \\
    \hline
    close & H5Gclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}DATASET} \\
    \hline
    create & H5Dcreate1 (deprecated) \\
           & H5Dcreate2 \\
    \hline
    open & H5Dopen1 (deprecated) \\
         & H5Dopen2 \\
    \hline
    read & H5Dread \\
    \hline
    write & H5Dwrite \\
    \hline
    get & H5Dget\_access\_plist \\
        & H5Dget\_create\_plist \\
        & H5Dextend \\
        & H5Dget\_offset \\
        & H5Dget\_space \\
        & H5Dget\_space\_status \\
        & H5Dget\_storage\_size \\
        & H5Dget\_type \\
    \hline
    specific & H5Dextend (deprecated) \\
             & H5Dflush \\
             & H5Drefresh \\
             & H5Dset\_extent \\
    \hline
    close & H5Dclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}OBJECT} \\
    \hline
    open & H5Oopen \\
         & H5Oopen\_by\_addr \\
         & H5Oopen\_by\_idx \\
         & H5Oopen\_by\_name \\
    \hline
    copy & H5Ocopy \\
    \hline
    get & N/A \\
    \hline
    specific & H5Odecr\_refcount \\
             & H5Oexists\_by\_name \\
             & H5Oflush \\
             & H5O\_incr\_refcount \\
             & H5Orefresh \\
             & H5Ovisit\_by\_name1 (deprecated) \\
             & H5Ovisit\_by\_name2 \\
             & H5Ovisit1 (deprecated) \\
             & H5Ovisit2 \\
    \hline
    close & H5Oclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}LINK} \\
    \hline
    create & H5Glink (deprecated) \\
           & H5Glink2 (deprecated) \\
           & H5Lcreate\_hard \\
           & H5Lcreate\_soft \\
           & H5Lcreate\_ud \\
           & H5Olink \\
    \hline
    copy & H5Lcopy \\
    \hline
    move & H5Gmove (deprecated) \\
         & H5Gmove2 (deprecated) \\
         & H5Lmove \\
    \hline
    get & H5Gget\_linkval (deprecated) \\
        & H5Lget\_info \\
        & H5Lget\_info\_by\_idx \\
        & H5Lget\_name\_by\_idx \\
        & H5Lget\_val \\
        & H5Lget\_val\_by\_idx \\
    \hline
    specific & H5Gunlink (deprecated) \\
             & H5Ldelete \\
             & H5Ldelete\_by\_idx \\
             & H5Lexists \\
             & H5Literate \\
             & H5Literate\_by\_name \\
             & H5Lvisit \\
             & H5Lvisit\_by\_name \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}DATATYPE} \\
    \hline
    commit & H5Tcommit1 (deprecated) \\
           & H5Tcommit2 \\
           & H5Tcommit\_anon \\
    \hline
    open & H5Topen1 (deprecated) \\
         & H5Topen2 \\
    \hline
    get & H5Tget\_create\_plist \\
    \hline
    specific & H5Tflush \\
             & H5Trefresh \\
    \hline
    close & H5Tclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}ATTRIBUTE} \\
    \hline
    create & H5Acreate1 (deprecated) \\
           & H5Acreate2 \\
           & H5Acreate\_by\_name \\
    \hline
    open & H5Aopen \\
         & H5Aopen\_by\_idx \\
         & H5Aopen\_by\_name \\
         & H5Aopen\_idx (deprecated)\\
         & H5Aopen\_name (deprecated) \\
    \hline
    read & H5Aread \\
    \hline
    write & H5Awrite \\
    \hline
    get & H5Aget\_get\_create\_plist \\
        & H5Aget\_info \\
        & H5Aget\_info\_by\_idx \\
        & H5Aget\_info\_by\_name \\
        & H5Aget\_name \\
        & H5Aget\_name\_by\_idx \\
        & H5Aget\_space \\
        & H5Aget\_storage\_size \\
        & H5Aget\_type \\
    \hline
    specific & H5Adelete \\
             & H5Adelete\_by\_idx \\
             & H5Adelete\_by\_name \\
             & H5Aexists \\
             & H5Aexists\_by\_name \\
             & H5Aiterate1 (deprecated) \\
             & H5Aiterate2 \\
             & H5Aiterate\_by\_name \\
             & H5Arename \\
             & H5Arename\_by\_name \\
    \hline
    close & H5Aclose \\
    \hline
\caption{Breakdown of HDF5 API calls by VOL callback}
\end{longtable}