### 2019-11-25 RFCs/HDF5/VOL/connector_author_guide/intro.tex

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

\section{Introduction}
The Virtual Object Layer (VOL) is an abstraction layer in the HDF5
library which intercepts all API calls that could potentially access
objects in an HDF5 container and forwards those calls to object drivers
referred to as \textit{VOL connectors}. The architecture of this feature
is described in the VOL User Guide and VOL Architecture and Internals
Documentation and will not be duplicated here.

This guide is for people who are interested in developing their own VOL
connector for the HDF5 library. It is assumed that the reader has good knowledge
of the VOL architecture obtained by reading the VOL architectural design
document~\cite{vol:rfc}.