### 2020-02-24 RFCs/HDF5/VOL/connector_author_guide/appendix.tex
Updated VOL user guide and connector author guide to be in agreement with final 1.12.0 changes.
Former-commit-id: f656be3383755816b47e431b1dabfeb7ee3e402a

\begin{appendices}

\pagebreak

\section{Mapping of VOL Callbacks to HDF5 API Calls}
\label{sec:A}

\begin{longtable}{ |l|>{\raggedright\arraybackslash}l| }
    \hline
    \multicolumn{1}{|c|}{VOL Callback} & \multicolumn{1}{|c|}{HDF5 API Call} \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}FILE} \\
    \hline
    create & H5Fcreate \\
    \hline
    open & H5Fopen \\
    \hline
    get & H5Fget\_access\_plist \\
        & H5Fget\_create\_plist \\
        & H5Fget\_fileno \\
        & H5Fget\_intent \\
        & H5Fget\_name \\
        & H5Fget\_obj\_count \\
        & H5Fget\_obj\_ids \\
    \hline
    specific & H5Fdelete \\
             & H5Fflush \\
             & H5Fis\_accessible \\
             & H5Fis\_hdf5 (deprecated, hard-coded to use native connector) \\
             & H5Fmount \\
             & H5Freopen \\
             & H5Funmount \\
    \hline
    close & H5Fclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}GROUP} \\
    \hline
    create & H5Gcreate1 (deprecated) \\
           & H5Gcreate2 \\
           & H5Gcreate\_anon \\
    \hline
    open & H5Gopen1 (deprecated) \\
         & H5Gopen2 \\
    \hline
    get & H5Gget\_create\_plist \\
        & H5Gget\_info \\
        & H5Gget\_info\_by\_idx \\
        & H5Gget\_info\_by\_name \\
        & H5Gget\_num\_objs (deprecated) \\
    \hline
    specific & H5Gflush \\
             & H5Grefresh \\
    \hline
    close & H5Gclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}DATASET} \\
    \hline
    create & H5Dcreate1 (deprecated) \\
           & H5Dcreate2 \\
    \hline
    open & H5Dopen1 (deprecated) \\
         & H5Dopen2 \\
    \hline
    read & H5Dread \\
    \hline
    write & H5Dwrite \\
    \hline
    get & H5Dget\_access\_plist \\
        & H5Dget\_create\_plist \\
        & H5Dextend \\
        & H5Dget\_offset \\
        & H5Dget\_space \\
        & H5Dget\_space\_status \\
        & H5Dget\_storage\_size \\
        & H5Dget\_type \\
    \hline
    specific & H5Dextend (deprecated) \\
             & H5Dflush \\
             & H5Drefresh \\
             & H5Dset\_extent \\
    \hline
    close & H5Dclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}OBJECT} \\
    \hline
    open & H5Oopen \\
         & H5Oopen\_by\_addr (deprecated) \\
         & H5Oopen\_by\_idx \\
         & H5Oopen\_by\_name \\
         & H5Oopen\_by\_token \\
    \hline
    copy & H5Ocopy \\
    \hline
    get & H5Oget\_info1 (deprecated) \\
        & H5Oget\_info2 (deprecated) \\
        & H5Oget\_info3 \\
    \hline
    specific & H5Odecr\_refcount \\
             & H5Oexists\_by\_name \\
             & H5Oflush \\
             & H5O\_incr\_refcount \\
             & H5Orefresh \\
             & H5Ovisit\_by\_name1 (deprecated) \\
             & H5Ovisit\_by\_name2 (deprecated) \\
             & H5Ovisit\_by\_name3 \\
             & H5Ovisit1 (deprecated) \\
             & H5Ovisit2 (deprecated) \\
             & H5Ovisit3 \\
    \hline
    close & H5Oclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}LINK} \\
    \hline
    create & H5Glink (deprecated) \\
           & H5Glink2 (deprecated) \\
           & H5Lcreate\_hard \\
           & H5Lcreate\_soft \\
           & H5Lcreate\_ud \\
           & H5Olink \\
    \hline
    copy & H5Lcopy \\
    \hline
    move & H5Gmove (deprecated) \\
         & H5Gmove2 (deprecated) \\
         & H5Lmove \\
    \hline
    get & H5Gget\_linkval (deprecated) \\
        & H5Lget\_info1 (deprecated) \\
        & H5Lget\_info2 \\
        & H5Lget\_info\_by\_idx \\
        & H5Lget\_name\_by\_idx \\
        & H5Lget\_val \\
        & H5Lget\_val\_by\_idx \\
    \hline
    specific & H5Gunlink (deprecated) \\
             & H5Ldelete \\
             & H5Ldelete\_by\_idx \\
             & H5Lexists \\
             & H5Literate1 (deprecated) \\
             & H5Literate2 \\
             & H5Literate\_by\_name1 (deprecated) \\
             & H5Literate\_by\_name2 \\
             & H5Lvisit1 (deprecated) \\
             & H5Lvisit2 \\
             & H5Lvisit\_by\_name1 (deprecated) \\
             & H5Lvisit\_by\_name2 \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}DATATYPE} \\
    \hline
    commit & H5Tcommit1 (deprecated) \\
           & H5Tcommit2 \\
           & H5Tcommit\_anon \\
    \hline
    open & H5Topen1 (deprecated) \\
         & H5Topen2 \\
    \hline
    get & H5Tget\_create\_plist \\
    \hline
    specific & H5Tflush \\
             & H5Trefresh \\
    \hline
    close & H5Tclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}ATTRIBUTE} \\
    \hline
    create & H5Acreate1 (deprecated) \\
           & H5Acreate2 \\
           & H5Acreate\_by\_name \\
    \hline
    open & H5Aopen \\
         & H5Aopen\_by\_idx \\
         & H5Aopen\_by\_name \\
         & H5Aopen\_idx (deprecated)\\
         & H5Aopen\_name (deprecated) \\
    \hline
    read & H5Aread \\
    \hline
    write & H5Awrite \\
    \hline
    get & H5Aget\_get\_create\_plist \\
        & H5Aget\_info \\
        & H5Aget\_info\_by\_idx \\
        & H5Aget\_info\_by\_name \\
        & H5Aget\_name \\
        & H5Aget\_name\_by\_idx \\
        & H5Aget\_space \\
        & H5Aget\_storage\_size \\
        & H5Aget\_type \\
    \hline
    specific & H5Adelete \\
             & H5Adelete\_by\_idx \\
             & H5Adelete\_by\_name \\
             & H5Aexists \\
             & H5Aexists\_by\_name \\
             & H5Aiterate1 (deprecated) \\
             & H5Aiterate2 \\
             & H5Aiterate\_by\_name \\
             & H5Arename \\
             & H5Arename\_by\_name \\
    \hline
    close & H5Aclose \\
    \hline
\caption{Breakdown of HDF5 API calls by VOL callback}
\end{longtable}

\pagebreak

\section{Callback Wrapper API Calls for Passthrough Connector Authors}
\label{sec:B}

From {\tt H5VL\_connector\_passthru.h}

\begin{lstlisting}

/* Helper routines for VOL connector authors */                                  
herr_t H5VLcmp_connector_cls(int *cmp, hid_t connector_id1, hid_t connector_id2);
hid_t H5VLwrap_register(void *obj, H5I_type_t type);                      
herr_t H5VLretrieve_lib_state(void **state);                              
herr_t H5VLrestore_lib_state(const void *state);                          
herr_t H5VLreset_lib_state(void);                                         
herr_t H5VLfree_lib_state(void *state);

/* Pass-through callbacks */                                                     
void *H5VLget_object(void *obj, hid_t connector_id);                      
herr_t H5VLget_wrap_ctx(void *obj, hid_t connector_id, void **wrap_ctx);  
void *H5VLwrap_object(void *obj, H5I_type_t obj_type, hid_t connector_id, void *wrap_ctx);                                                             
void *H5VLunwrap_object(void *obj, hid_t connector_id);                   
herr_t H5VLfree_wrap_ctx(void *wrap_ctx, hid_t connector_id);

/* Public wrappers for generic callbacks */                                      
herr_t H5VLinitialize(hid_t connector_id, hid_t vipl_id);                 
herr_t H5VLterminate(hid_t connector_id);                                 
herr_t H5VLget_cap_flags(hid_t connector_id, unsigned *cap_flags);        
herr_t H5VLget_value(hid_t connector_id, H5VL_class_value_t *conn_value);

/* Public wrappers for info fields and callbacks */                              
herr_t H5VLcopy_connector_info(hid_t connector_id, void **dst_vol_info, void *src_vol_info);
herr_t H5VLcmp_connector_info(int *cmp, hid_t connector_id, const void *info1,
    const void *info2);                                                          
herr_t H5VLfree_connector_info(hid_t connector_id, void *vol_info);       
herr_t H5VLconnector_info_to_str(const void *info, hid_t connector_id, char **str);
herr_t H5VLconnector_str_to_info(const char *str, hid_t connector_id, void **info);

/* Public wrappers for attribute callbacks */                                    
void *H5VLattr_create(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *attr_name, hid_t type_id, hid_t space_id, hid_t acpl_id, hid_t aapl_id, hid_t dxpl_id, void **req);
void *H5VLattr_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t aapl_id, hid_t dxpl_id, void **req);
herr_t H5VLattr_read(void *attr, hid_t connector_id, hid_t dtype_id, void *buf, hid_t dxpl_id, void **req);
herr_t H5VLattr_write(void *attr, hid_t connector_id, hid_t dtype_id, const void *buf, hid_t dxpl_id, void **req);
herr_t H5VLattr_get(void *obj, hid_t connector_id, H5VL_attr_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLattr_specific(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, H5VL_attr_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLattr_optional(void *obj, hid_t connector_id, H5VL_attribute_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLattr_close(void *attr, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for dataset callbacks */                                      
void *H5VLdataset_create(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t lcpl_id, hid_t type_id, hid_t space_id, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req);
void *H5VLdataset_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
herr_t H5VLdataset_read(void *dset, hid_t connector_id, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t plist_id, void *buf, void **req);
herr_t H5VLdataset_write(void *dset, hid_t connector_id, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t plist_id, const void *buf, void **req);
herr_t H5VLdataset_get(void *dset, hid_t connector_id, H5VL_dataset_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdataset_specific(void *obj, hid_t connector_id, H5VL_dataset_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdataset_optional(void *obj, hid_t connector_id, H5VL_dataset_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdataset_close(void *dset, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for file callbacks */                                         
void *H5VLfile_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id, void **req);
void *H5VLfile_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req);
herr_t H5VLfile_get(void *file, hid_t connector_id, H5VL_file_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLfile_specific(void *obj, hid_t connector_id, H5VL_file_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLfile_optional(void *obj, hid_t connector_id, H5VL_file_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLfile_close(void *file, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for group callbacks */                                        
void *H5VLgroup_create(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t lcpl_id, hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id, void **req);
void *H5VLgroup_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t gapl_id, hid_t dxpl_id, void **req);
herr_t H5VLgroup_get(void *obj, hid_t connector_id, H5VL_group_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLgroup_specific(void *obj, hid_t connector_id, H5VL_group_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLgroup_optional(void *obj, hid_t connector_id, H5VL_group_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLgroup_close(void *grp, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for link callbacks */
herr_t H5VLlink_create(H5VL_link_create_type_t create_type, void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLlink_copy(void *src_obj, const H5VL_loc_params_t *loc_params1, void *dst_obj, const H5VL_loc_params_t *loc_params2, hid_t connector_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_move(void *src_obj, const H5VL_loc_params_t *loc_params1, void *dst_obj, const H5VL_loc_params_t *loc_params2, hid_t connector_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_get(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, H5VL_link_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLlink_specific(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, H5VL_link_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLlink_optional(void *obj, hid_t connector_id, H5VL_link_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);

/* Public wrappers for object callbacks */
void *H5VLobject_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, H5I_type_t *opened_type, hid_t dxpl_id, void **req);
herr_t H5VLobject_copy(void *src_obj, const H5VL_loc_params_t *loc_params1, const char *src_name, void *dst_obj, const H5VL_loc_params_t *loc_params2, const char *dst_name, hid_t connector_id, hid_t ocpypl_id, hid_t lcpl_id, hid_t dxpl_id, void **req);
herr_t H5VLobject_get(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, H5VL_object_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLobject_specific(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, H5VL_object_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLobject_optional(void *obj, hid_t connector_id, H5VL_object_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);

/* Public wrappers for named datatype callbacks */                               
void *H5VLdatatype_commit(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id, hid_t dxpl_id, void **req);
void *H5VLdatatype_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id, const char *name, hid_t tapl_id, hid_t dxpl_id, void **req);
herr_t H5VLdatatype_get(void *dt, hid_t connector_id, H5VL_datatype_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdatatype_specific(void *obj, hid_t connector_id, H5VL_datatype_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdatatype_optional(void *obj, hid_t connector_id, H5VL_datatype_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
herr_t H5VLdatatype_close(void *dt, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for connector/container introspection callbacks */            
H5_DLL herr_t H5VLintrospect_get_conn_cls(void *obj, hid_t connector_id, H5VL_get_conn_lvl_t lvl, const H5VL_class_t **conn_cls);                     
H5_DLL herr_t H5VLintrospect_opt_query(void *obj, hid_t connector_id, H5VL_subclass_t subcls, int opt_type, hbool_t *supported);

/* Public wrappers for asynchronous request callbacks */                         
herr_t H5VLrequest_wait(void *req, hid_t connector_id, uint64_t timeout, H5ES_status_t *status);
herr_t H5VLrequest_notify(void *req, hid_t connector_id, H5VL_request_notify_t cb, void *ctx);
herr_t H5VLrequest_cancel(void *req, hid_t connector_id);                 
herr_t H5VLrequest_specific(void *req, hid_t connector_id, H5VL_request_specific_t specific_type, va_list arguments);
herr_t H5VLrequest_optional(void *req, hid_t connector_id, H5VL_request_optional_t opt_type, va_list arguments);
herr_t H5VLrequest_free(void *req, hid_t connector_id);

/* Public wrappers for blob callbacks */                                         
herr_t H5VLblob_put(void *obj, hid_t connector_id, const void *buf, size_t size, void *blob_id, void *ctx);
herr_t H5VLblob_get(void *obj, hid_t connector_id, const void *blob_id, void *buf, size_t size, void *ctx);
herr_t H5VLblob_specific(void *obj, hid_t connector_id, void *blob_id, H5VL_blob_specific_t specific_type, va_list arguments);
H5_DLL herr_t H5VLblob_optional(void *obj, hid_t connector_id, void *blob_id, H5VL_blob_optional_t opt_type, va_list arguments);

/* Public wrappers for token callbacks */                                        
H5_DLL herr_t H5VLtoken_cmp(void *obj, hid_t connector_id, const H5O_token_t *token1, const H5O_token_t *token2, int *cmp_value);
H5_DLL herr_t H5VLtoken_to_str(void *obj, H5I_type_t obj_type, hid_t connector_id, const H5O_token_t *token, char **token_str);
H5_DLL herr_t H5VLtoken_from_str(void *obj, H5I_type_t obj_type, hid_t connector_id, const char *token_str, H5O_token_t *token);

/* Public wrappers for generic 'optional' callback */                            
herr_t H5VLoptional(void *obj, hid_t connector_id, int op_type, hid_t dxpl_id, void **req, va_list arguments);

\end{lstlisting}

\end{appendices}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: