### 2022-05-27 RFCs/HDF5/VOL/connector_author_guide/appendix.tex

\begin{appendices}

\pagebreak

\section{Mapping of VOL Callbacks to HDF5 API Calls}
\label{sec:A}

\begin{longtable}{ |l|>{\raggedright\arraybackslash}l| }
    \hline
    \multicolumn{1}{|c|}{VOL Callback} & \multicolumn{1}{|c|}{HDF5 API Call} \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}FILE} \\
    \hline
    create & H5Fcreate \\
    \hline
    open & H5Fopen \\
    \hline
    get & H5Fget\_access\_plist \\
        & H5Fget\_create\_plist \\
        & H5Fget\_fileno \\
        & H5Fget\_intent \\
        & H5Fget\_name \\
        & H5Fget\_obj\_count \\
        & H5Fget\_obj\_ids \\
    \hline
    specific & H5Fdelete \\
             & H5Fflush \\
             & H5Fis\_accessible \\
             & H5Fis\_hdf5 (deprecated, hard-coded to use native connector) \\
             & H5Freopen \\
    \hline
    close & H5Fclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}GROUP} \\
    \hline
    create & H5Gcreate1 (deprecated) \\
           & H5Gcreate2 \\
           & H5Gcreate\_anon \\
    \hline
    open & H5Gopen1 (deprecated) \\
         & H5Gopen2 \\
    \hline
    get & H5Gget\_create\_plist \\
        & H5Gget\_info \\
        & H5Gget\_info\_by\_idx \\
        & H5Gget\_info\_by\_name \\
        & H5Gget\_num\_objs (deprecated) \\
    \hline
    specific & H5Fmount \\
             & H5Funmount \\
             & H5Gflush \\
             & H5Grefresh \\
    \hline
    close & H5Gclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}DATASET} \\
    \hline
    create & H5Dcreate1 (deprecated) \\
           & H5Dcreate2 \\
    \hline
    open & H5Dopen1 (deprecated) \\
         & H5Dopen2 \\
    \hline
    read & H5Dread \\
    \hline
    write & H5Dwrite \\
    \hline
    get & H5Dget\_access\_plist \\
        & H5Dget\_create\_plist \\
        & H5Dextend \\
        & H5Dget\_space \\
        & H5Dget\_space\_status \\
        & H5Dget\_storage\_size \\
        & H5Dget\_type \\
    \hline
    specific & H5Dextend (deprecated) \\
             & H5Dflush \\
             & H5Drefresh \\
             & H5Dset\_extent \\
    \hline
    close & H5Dclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}OBJECT} \\
    \hline
    open & H5Oopen \\
         & H5Oopen\_by\_addr (deprecated) \\
         & H5Oopen\_by\_idx \\
         & H5Oopen\_by\_name \\
         & H5Oopen\_by\_token \\
    \hline
    copy & H5Ocopy \\
    \hline
    get & H5Oget\_info1 (deprecated) \\
        & H5Oget\_info2 (deprecated) \\
        & H5Oget\_info3 \\
    \hline
    specific & H5Odecr\_refcount \\
             & H5Oexists\_by\_name \\
             & H5Oflush \\
             & H5O\_incr\_refcount \\
             & H5Orefresh \\
             & H5Ovisit\_by\_name1 (deprecated) \\
             & H5Ovisit\_by\_name2 (deprecated) \\
             & H5Ovisit\_by\_name3 \\
             & H5Ovisit1 (deprecated) \\
             & H5Ovisit2 (deprecated) \\
             & H5Ovisit3 \\
    \hline
    close & H5Oclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}LINK} \\
    \hline
    create & H5Glink (deprecated) \\
           & H5Glink2 (deprecated) \\
           & H5Lcreate\_hard \\
           & H5Lcreate\_soft \\
           & H5Lcreate\_ud \\
           & H5Olink \\
    \hline
    copy & H5Lcopy \\
    \hline
    move & H5Gmove (deprecated) \\
         & H5Gmove2 (deprecated) \\
         & H5Lmove \\
    \hline
    get & H5Gget\_linkval (deprecated) \\
        & H5Lget\_info1 (deprecated) \\
        & H5Lget\_info2 \\
        & H5Lget\_info\_by\_idx \\
        & H5Lget\_name\_by\_idx \\
        & H5Lget\_val \\
        & H5Lget\_val\_by\_idx \\
    \hline
    specific & H5Gunlink (deprecated) \\
             & H5Ldelete \\
             & H5Ldelete\_by\_idx \\
             & H5Lexists \\
             & H5Literate1 (deprecated) \\
             & H5Literate2 \\
             & H5Literate\_by\_name1 (deprecated) \\
             & H5Literate\_by\_name2 \\
             & H5Lvisit1 (deprecated) \\
             & H5Lvisit2 \\
             & H5Lvisit\_by\_name1 (deprecated) \\
             & H5Lvisit\_by\_name2 \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}DATATYPE} \\
    \hline
    commit & H5Tcommit1 (deprecated) \\
           & H5Tcommit2 \\
           & H5Tcommit\_anon \\
    \hline
    open & H5Topen1 (deprecated) \\
         & H5Topen2 \\
    \hline
    get & H5Tget\_create\_plist \\
    \hline
    specific & H5Tflush \\
             & H5Trefresh \\
    \hline
    close & H5Tclose \\
    \hline
    \multicolumn{2}{|>{\raggedright\arraybackslash}p{\linewidth}|}{\cellcolor{gray!35}ATTRIBUTE} \\
    \hline
    create & H5Acreate1 (deprecated) \\
           & H5Acreate2 \\
           & H5Acreate\_by\_name \\
    \hline
    open & H5Aopen \\
         & H5Aopen\_by\_idx \\
         & H5Aopen\_by\_name \\
         & H5Aopen\_idx (deprecated)\\
         & H5Aopen\_name (deprecated) \\
    \hline
    read & H5Aread \\
    \hline
    write & H5Awrite \\
    \hline
    get & H5Aget\_get\_create\_plist \\
        & H5Aget\_info \\
        & H5Aget\_info\_by\_idx \\
        & H5Aget\_info\_by\_name \\
        & H5Aget\_name \\
        & H5Aget\_name\_by\_idx \\
        & H5Aget\_space \\
        & H5Aget\_storage\_size \\
        & H5Aget\_type \\
    \hline
    specific & H5Adelete \\
             & H5Adelete\_by\_idx \\
             & H5Adelete\_by\_name \\
             & H5Aexists \\
             & H5Aexists\_by\_name \\
             & H5Aiterate1 (deprecated) \\
             & H5Aiterate2 \\
             & H5Aiterate\_by\_name \\
             & H5Arename \\
             & H5Arename\_by\_name \\
    \hline
    close & H5Aclose \\
    \hline
\caption{Breakdown of HDF5 API calls by VOL callback}
\end{longtable}

\pagebreak

\section{Callback Wrapper API Calls for Passthrough Connector Authors}
\label{sec:B}

From {\tt H5VL\_connector\_passthru.h}

\begin{lstlisting}

/* Pass-through callbacks */
void * H5VLget_object(void *obj, hid_t connector_id);
herr_t H5VLget_wrap_ctx(void *obj, hid_t connector_id, void **wrap_ctx);
void * H5VLwrap_object(void *obj, H5I_type_t obj_type, hid_t connector_id, void *wrap_ctx);
void * H5VLunwrap_object(void *obj, hid_t connector_id);
herr_t H5VLfree_wrap_ctx(void *wrap_ctx, hid_t connector_id);

/* Public wrappers for generic callbacks */
herr_t H5VLinitialize(hid_t connector_id, hid_t vipl_id);
herr_t H5VLterminate(hid_t connector_id);
herr_t H5VLget_cap_flags(hid_t connector_id, unsigned *cap_flags);
herr_t H5VLget_value(hid_t connector_id, H5VL_class_value_t *conn_value);

/* Public wrappers for info fields and callbacks */
herr_t H5VLcopy_connector_info(hid_t connector_id, void **dst_vol_info, void *src_vol_info);
herr_t H5VLcmp_connector_info(int *cmp, hid_t connector_id, const void *info1, const void *info2);
herr_t H5VLfree_connector_info(hid_t connector_id, void *vol_info);
herr_t H5VLconnector_info_to_str(const void *info, hid_t connector_id, char **str);
herr_t H5VLconnector_str_to_info(const char *str, hid_t connector_id, void **info);

/* Public wrappers for attribute callbacks */
void * H5VLattr_create(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                              const char *attr_name, hid_t type_id, hid_t space_id, hid_t acpl_id,
                              hid_t aapl_id, hid_t dxpl_id, void **req);
void * H5VLattr_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                            const char *name, hid_t aapl_id, hid_t dxpl_id, void **req);
herr_t H5VLattr_read(void *attr, hid_t connector_id, hid_t dtype_id, void *buf, hid_t dxpl_id,
                            void **req);
herr_t H5VLattr_write(void *attr, hid_t connector_id, hid_t dtype_id, const void *buf, hid_t dxpl_id,
                             void **req);
herr_t H5VLattr_get(void *obj, hid_t connector_id, H5VL_attr_get_args_t *args, hid_t dxpl_id,
                           void **req);
herr_t H5VLattr_specific(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                H5VL_attr_specific_args_t *args, hid_t dxpl_id, void **req);
herr_t H5VLattr_optional(void *obj, hid_t connector_id, H5VL_optional_args_t *args, hid_t dxpl_id,
                                void **req);
herr_t H5VLattr_close(void *attr, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for dataset callbacks */
void * H5VLdataset_create(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                 const char *name, hid_t lcpl_id, hid_t type_id, hid_t space_id, hid_t dcpl_id,
                                 hid_t dapl_id, hid_t dxpl_id, void **req);
void * H5VLdataset_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                               const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
herr_t H5VLdataset_read(void *dset, hid_t connector_id, hid_t mem_type_id, hid_t mem_space_id,
                               hid_t file_space_id, hid_t plist_id, void *buf, void **req);
herr_t H5VLdataset_write(void *dset, hid_t connector_id, hid_t mem_type_id, hid_t mem_space_id,
                                hid_t file_space_id, hid_t plist_id, const void *buf, void **req);
herr_t H5VLdataset_get(void *dset, hid_t connector_id, H5VL_dataset_get_args_t *args, hid_t dxpl_id,
                              void **req);
herr_t H5VLdataset_specific(void *obj, hid_t connector_id, H5VL_dataset_specific_args_t *args,
                                   hid_t dxpl_id, void **req);
herr_t H5VLdataset_optional(void *obj, hid_t connector_id, H5VL_optional_args_t *args, hid_t dxpl_id,
                                   void **req);
herr_t H5VLdataset_close(void *dset, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for named datatype callbacks */
void * H5VLdatatype_commit(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                  const char *name, hid_t type_id, hid_t lcpl_id, hid_t tcpl_id, hid_t tapl_id,
                                  hid_t dxpl_id, void **req);
void * H5VLdatatype_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                const char *name, hid_t tapl_id, hid_t dxpl_id, void **req);
herr_t H5VLdatatype_get(void *dt, hid_t connector_id, H5VL_datatype_get_args_t *args, hid_t dxpl_id,
                               void **req);
herr_t H5VLdatatype_specific(void *obj, hid_t connector_id, H5VL_datatype_specific_args_t *args,
                                    hid_t dxpl_id, void **req);
herr_t H5VLdatatype_optional(void *obj, hid_t connector_id, H5VL_optional_args_t *args, hid_t dxpl_id,
                                    void **req);
herr_t H5VLdatatype_close(void *dt, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for file callbacks */
void * H5VLfile_create(const char *name, unsigned flags, hid_t fcpl_id, hid_t fapl_id, hid_t dxpl_id,
                              void **req);
void * H5VLfile_open(const char *name, unsigned flags, hid_t fapl_id, hid_t dxpl_id, void **req);
herr_t H5VLfile_get(void *file, hid_t connector_id, H5VL_file_get_args_t *args, hid_t dxpl_id,
                           void **req);
herr_t H5VLfile_specific(void *obj, hid_t connector_id, H5VL_file_specific_args_t *args, hid_t dxpl_id,
                                void **req);
herr_t H5VLfile_optional(void *obj, hid_t connector_id, H5VL_optional_args_t *args, hid_t dxpl_id,
                                void **req);
herr_t H5VLfile_close(void *file, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for group callbacks */
void * H5VLgroup_create(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                               const char *name, hid_t lcpl_id, hid_t gcpl_id, hid_t gapl_id, hid_t dxpl_id,
                               void **req);
void * H5VLgroup_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                             const char *name, hid_t gapl_id, hid_t dxpl_id, void **req);
herr_t H5VLgroup_get(void *obj, hid_t connector_id, H5VL_group_get_args_t *args, hid_t dxpl_id,
                            void **req);
herr_t H5VLgroup_specific(void *obj, hid_t connector_id, H5VL_group_specific_args_t *args,
                                 hid_t dxpl_id, void **req);
herr_t H5VLgroup_optional(void *obj, hid_t connector_id, H5VL_optional_args_t *args, hid_t dxpl_id,
                                 void **req);
herr_t H5VLgroup_close(void *grp, hid_t connector_id, hid_t dxpl_id, void **req);

/* Public wrappers for link callbacks */
herr_t H5VLlink_create(H5VL_link_create_args_t *args, void *obj, const H5VL_loc_params_t *loc_params,
                              hid_t connector_id, hid_t lcpl_id, hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_copy(void *src_obj, const H5VL_loc_params_t *loc_params1, void *dst_obj,
                            const H5VL_loc_params_t *loc_params2, hid_t connector_id, hid_t lcpl_id,
                            hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_move(void *src_obj, const H5VL_loc_params_t *loc_params1, void *dst_obj,
                            const H5VL_loc_params_t *loc_params2, hid_t connector_id, hid_t lcpl_id,
                            hid_t lapl_id, hid_t dxpl_id, void **req);
herr_t H5VLlink_get(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                           H5VL_link_get_args_t *args, hid_t dxpl_id, void **req);
herr_t H5VLlink_specific(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                H5VL_link_specific_args_t *args, hid_t dxpl_id, void **req);
herr_t H5VLlink_optional(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                H5VL_optional_args_t *args, hid_t dxpl_id, void **req);

/* Public wrappers for object callbacks */
void * H5VLobject_open(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                              H5I_type_t *opened_type, hid_t dxpl_id, void **req);
herr_t H5VLobject_copy(void *src_obj, const H5VL_loc_params_t *loc_params1, const char *src_name,
                              void *dst_obj, const H5VL_loc_params_t *loc_params2, const char *dst_name,
                              hid_t connector_id, hid_t ocpypl_id, hid_t lcpl_id, hid_t dxpl_id, void **req);
herr_t H5VLobject_get(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                             H5VL_object_get_args_t *args, hid_t dxpl_id, void **req);
herr_t H5VLobject_specific(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                  H5VL_object_specific_args_t *args, hid_t dxpl_id, void **req);
herr_t H5VLobject_optional(void *obj, const H5VL_loc_params_t *loc_params, hid_t connector_id,
                                  H5VL_optional_args_t *args, hid_t dxpl_id, void **req);

/* Public wrappers for connector/container introspection callbacks */
herr_t H5VLintrospect_get_conn_cls(void *obj, hid_t connector_id, H5VL_get_conn_lvl_t lvl,
                                          const H5VL_class_t **conn_cls);
herr_t H5VLintrospect_get_cap_flags(const void *info, hid_t connector_id, unsigned *cap_flags);
herr_t H5VLintrospect_opt_query(void *obj, hid_t connector_id, H5VL_subclass_t subcls, int opt_type,
                                       uint64_t *flags);

/* Public wrappers for asynchronous request callbacks */
herr_t H5VLrequest_wait(void *req, hid_t connector_id, uint64_t timeout,
                               H5VL_request_status_t *status);
herr_t H5VLrequest_notify(void *req, hid_t connector_id, H5VL_request_notify_t cb, void *ctx);
herr_t H5VLrequest_cancel(void *req, hid_t connector_id, H5VL_request_status_t *status);
herr_t H5VLrequest_specific(void *req, hid_t connector_id, H5VL_request_specific_args_t *args);
herr_t H5VLrequest_optional(void *req, hid_t connector_id, H5VL_optional_args_t *args);
herr_t H5VLrequest_free(void *req, hid_t connector_id);

/* Public wrappers for blob callbacks */
herr_t H5VLblob_put(void *obj, hid_t connector_id, const void *buf, size_t size, void *blob_id,
                           void *ctx);
herr_t H5VLblob_get(void *obj, hid_t connector_id, const void *blob_id, void *buf, size_t size,
                           void *ctx);
herr_t H5VLblob_specific(void *obj, hid_t connector_id, void *blob_id,
                                H5VL_blob_specific_args_t *args);
herr_t H5VLblob_optional(void *obj, hid_t connector_id, void *blob_id, H5VL_optional_args_t *args);

/* Public wrappers for token callbacks */
herr_t H5VLtoken_cmp(void *obj, hid_t connector_id, const H5O_token_t *token1,
                            const H5O_token_t *token2, int *cmp_value);
herr_t H5VLtoken_to_str(void *obj, H5I_type_t obj_type, hid_t connector_id, const H5O_token_t *token,
                               char **token_str);
herr_t H5VLtoken_from_str(void *obj, H5I_type_t obj_type, hid_t connector_id, const char *token_str,
                                 H5O_token_t *token);

/* Public wrappers for generic 'optional' callback */
herr_t H5VLoptional(void *obj, hid_t connector_id, H5VL_optional_args_t *args, hid_t dxpl_id,
                           void **req);
\end{lstlisting}

\pagebreak

\section{Native VOL Connector Optional Values By Subclass}
\label{sec:C}

\begin{lstlisting}
/* H5VL_SUBCLS_ATTR */
/* H5Aiterate (deprecated routine) */
#define H5VL_NATIVE_ATTR_ITERATE_OLD 0

/* H5VL_SUBCLS_DATASET */
/* H5Dformat_convert (internal) */
#define H5VL_NATIVE_DATASET_FORMAT_CONVERT          0  
/* H5Dget_chunk_index_type */
#define H5VL_NATIVE_DATASET_GET_CHUNK_INDEX_TYPE    1  
/* H5Dget_chunk_storage_size */
#define H5VL_NATIVE_DATASET_GET_CHUNK_STORAGE_SIZE  2  
/* H5Dget_num_chunks */
#define H5VL_NATIVE_DATASET_GET_NUM_CHUNKS          3  
/* H5Dget_chunk_info */
#define H5VL_NATIVE_DATASET_GET_CHUNK_INFO_BY_IDX   4  
/* H5Dget_chunk_info_by_coord */
#define H5VL_NATIVE_DATASET_GET_CHUNK_INFO_BY_COORD 5  
/* H5Dchunk_read */
#define H5VL_NATIVE_DATASET_CHUNK_READ              6  
/* H5Dchunk_write */
#define H5VL_NATIVE_DATASET_CHUNK_WRITE             7  
/* H5Dvlen_get_buf_size */
#define H5VL_NATIVE_DATASET_GET_VLEN_BUF_SIZE       8  
/* H5Dget_offset */
#define H5VL_NATIVE_DATASET_GET_OFFSET              9  
/* H5Dget_offset */
#define H5VL_NATIVE_DATASET_CHUNK_ITER              10

/* H5VL_SUBCLS_FILE */
/* H5Fclear_elink_file_cache */
#define H5VL_NATIVE_FILE_CLEAR_ELINK_CACHE            0  
/* H5Fget_file_image */
#define H5VL_NATIVE_FILE_GET_FILE_IMAGE               1  
/* H5Fget_free_sections */
#define H5VL_NATIVE_FILE_GET_FREE_SECTIONS            2  
/* H5Fget_freespace */
#define H5VL_NATIVE_FILE_GET_FREE_SPACE               3  
/* H5Fget_info1/2 */
#define H5VL_NATIVE_FILE_GET_INFO                     4  
/* H5Fget_mdc_config */
#define H5VL_NATIVE_FILE_GET_MDC_CONF                 5  
/* H5Fget_mdc_hit_rate */
#define H5VL_NATIVE_FILE_GET_MDC_HR                   6  
/* H5Fget_mdc_size */
#define H5VL_NATIVE_FILE_GET_MDC_SIZE                 7  
/* H5Fget_filesize */
#define H5VL_NATIVE_FILE_GET_SIZE                     8  
/* H5Fget_vfd_handle */
#define H5VL_NATIVE_FILE_GET_VFD_HANDLE               9  
/* H5Freset_mdc_hit_rate_stats */
#define H5VL_NATIVE_FILE_RESET_MDC_HIT_RATE           10 
/* H5Fset_mdc_config */
#define H5VL_NATIVE_FILE_SET_MDC_CONFIG               11 
/* H5Fget_metadata_read_retry_info */
#define H5VL_NATIVE_FILE_GET_METADATA_READ_RETRY_INFO 12 
/* H5Fstart_swmr_write */
#define H5VL_NATIVE_FILE_START_SWMR_WRITE             13 
/* H5Fstart_mdc_logging */
#define H5VL_NATIVE_FILE_START_MDC_LOGGING            14 
/* H5Fstop_mdc_logging */
#define H5VL_NATIVE_FILE_STOP_MDC_LOGGING             15 
/* H5Fget_mdc_logging_status */
#define H5VL_NATIVE_FILE_GET_MDC_LOGGING_STATUS       16 
/* H5Fformat_convert */
#define H5VL_NATIVE_FILE_FORMAT_CONVERT               17 
/* H5Freset_page_buffering_stats */
#define H5VL_NATIVE_FILE_RESET_PAGE_BUFFERING_STATS   18 
/* H5Fget_page_buffering_stats */
#define H5VL_NATIVE_FILE_GET_PAGE_BUFFERING_STATS     19 
/* H5Fget_mdc_image_info */
#define H5VL_NATIVE_FILE_GET_MDC_IMAGE_INFO           20 
/* H5Fget_eoa */
#define H5VL_NATIVE_FILE_GET_EOA                      21 
/* H5Fincrement_filesize */
#define H5VL_NATIVE_FILE_INCR_FILESIZE                22 
/* H5Fset_latest_format/libver_bounds */
#define H5VL_NATIVE_FILE_SET_LIBVER_BOUNDS            23 
/* H5Fget_dset_no_attrs_hint */
#define H5VL_NATIVE_FILE_GET_MIN_DSET_OHDR_FLAG       24 
/* H5Fset_dset_no_attrs_hint */
#define H5VL_NATIVE_FILE_SET_MIN_DSET_OHDR_FLAG       25 
/* H5Fget_mpi_atomicity */
#define H5VL_NATIVE_FILE_GET_MPI_ATOMICITY 26 
/* H5Fset_mpi_atomicity */
#define H5VL_NATIVE_FILE_SET_MPI_ATOMICITY 27 
/* Adjust file after open, with wrapping context */
#define H5VL_NATIVE_FILE_POST_OPEN 28

/* H5VL_SUBCLS_GROUP */
/* HG5Giterate (deprecated routine) */
#define H5VL_NATIVE_GROUP_ITERATE_OLD 0 
/* HG5Gget_objinfo (deprecated routine) */
#define H5VL_NATIVE_GROUP_GET_OBJINFO 1

/* H5VL_SUBCLS_OBJECT */
/* H5G|H5Oget_comment, H5Oget_comment_by_name */
#define H5VL_NATIVE_OBJECT_GET_COMMENT              0 
/* H5G|H5Oset_comment, H5Oset_comment_by_name */
#define H5VL_NATIVE_OBJECT_SET_COMMENT              1 
/* H5Odisable_mdc_flushes */
#define H5VL_NATIVE_OBJECT_DISABLE_MDC_FLUSHES      2 
/* H5Oenable_mdc_flushes */
#define H5VL_NATIVE_OBJECT_ENABLE_MDC_FLUSHES       3 
/* H5Oare_mdc_flushes_disabled */
#define H5VL_NATIVE_OBJECT_ARE_MDC_FLUSHES_DISABLED 4 
/* H5Oget_native_info(_by_idx, _by_name) */
#define H5VL_NATIVE_OBJECT_GET_NATIVE_INFO          5 
\end{lstlisting}

\end{appendices}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: