### 2023-01-27 RFCs/HDF5_Library/VFD_Subfiling/user_guide/sections/acknowledgements.tex

\documentclass[../main.tex]{subfiles}

\begin{document}

\newpage

\section*{Acknowledgments}
\label{acknowledgements}

% Include this section in the table of contents
\addcontentsline{toc}{section}{Acknowledgements}
This work was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.

\end{document}