### 2023-01-27 RFCs/HDF5_Library/VFD_Subfiling/user_guide/sections/limitations.tex

\documentclass[../main.tex]{subfiles}

\begin{document}

\newpage

\section{Known problems and limitations}

\begin{itemize}
  \item Collective I/O is currently not supported, except for collective metadata
        reads/writes, as enabled by \texttt{H5Pset\_all\_coll\_metadata\_ops} and
        \texttt{H5Pset\_coll\_metadata\_write}

  \item The VFD should currently not be used with an HDF5 library that has been
        built with thread-safety enabled. This can cause deadlocks when failures
        occur due to interactions between the VFD's internal threads and HDF5's
        global lock.

  \item The internal I/O task queue doesn't currently have any sophisticated error handling
        method, so errors are only reported at a course-grained level if an I/O request
        fails at the lowest level (pread/pwrite/etc.)
\end{itemize}

\end{document}