### 2023-01-27 RFCs/HDF5_Library/VFD_Subfiling/user_guide/sections/tips.tex

\documentclass[../main.tex]{subfiles}

\begin{document}

\newpage

\section{Tips and best practices}

\begin{itemize}
\item While the default configuration for the Subfiling VFD has been found
to be generally performant across different machines, it is still a good
idea to initially test a range of Subfiling configurations to see if there
is a particular one that generally performs best on the machine of interest.

\item When using a parallel file system that stripes data across different
storage targets (e.g., Lustre, GPFS, etc.), it is generally a good idea to
choose a Subfiling stripe size that is a multiple of the size of the data
stripes. For the best performance, it is also generally recommended that
the HDF5 application use the \texttt{H5Pset\_alignment} API routine to set
the alignment of objects in the file to match the Subfiling stripe size or
to at least be a multiple of the size of the file system's data stripes.
However, setting the alignment of objects in the file can potentially waste
space and vastly increase the size of the resulting file, so caution may be
needed when trying to find a good balance with this suggestion.

\item While not always leading to an improvement in performance, it can
sometimes be useful to enable collective metadata writes and, if the application's
access pattern allows for it, collective metadata reads in a parallel HDF5
application through use of the \texttt{H5Pset\_coll\_metadata\_write} and
\texttt{H5Pset\_all\_coll\_metadata\_ops} API routines.

\item When measuring the performance of the Subfiling feature, be aware that
results may sometimes be artificially high due to file system caching.
Since all I/O to a particular subfile is directed to a single I/O Concentrator
on a machine node, there are many opportunities for data from subfiles to
be served from or written to file system caching layers.
\end{itemize}

\end{document}