### 2023-01-27 RFCs/HDF5_Library/VFD_Subfiling/user_guide/support/glossary.tex

\makeglossaries

% Glossary terms 
\newglossaryentry{hid}
{
        name=HDF5 Identifier (HID),
        description={An integer which the library (H5I in HDF5) associates with an
                     "object" in memory, such as a file, FAPL, dataset, etc.}
}

\newglossaryentry{FAPL}
{
        name=File Access Property List (FAPL),
        description={An HDF5 \href{https://docs.hdfgroup.org/hdf5/develop/group___h5_p.html}{Property List}
                     object that is used to control how a file is accessed.}
}

\newglossaryentry{VFL}
{
        name=Virtual File Layer (VFL),
        description={A logical layer in the HDF5 library responsible for isolating the
                     library's operations from the underlying storage mechanism
                     (filesystem, e.g.).}
}

\newglossaryentry{VFD}
{
        name=Virtual File Driver (VFD),
        description={An implementation of a Virtual File Driver "class" – a structure
                     containing function pointers and other data required for Virtual File
                     operations.}
}

\newglossaryentry{tVFD}
{
        name=Terminal VFD,
        description={VFD which performs operations on storage and does not support
                     any child VFD as part of its own operation.}
}

\newglossaryentry{pVFD}
{
        name=Passthrough VFD,
        description={VFD which redirects operations to other “child” VFDs.}
}

\newglossaryentry{cVFD}
{
        name=Child VFD,
        description={VFD which receives direction from a “passthrough” VFD, rather
                     than from the HDF5 library directly.}
}

\newglossaryentry{hdf5_plugin}
{
        name=Plugin,
        description={Binary code which can be loaded and utilized at runtime. Extant
                     HDF5 examples include filter and VOL connector plugins.}
}

% Glossary acronyms 
\newacronym{hdf5}{HDF5}{Hierarchical Data Format, v5}

\newacronym{rfc}{RFC}{Request for Comments}