### 2023-06-20 RFCs/HDF5_Library/VFD_Subfiling/user_guide/sections/acknowledgements.tex
Update Subfiling User's Guide with notes about stripe count

\documentclass[../main.tex]{subfiles}

\begin{document}

\newpage

\section*{Acknowledgments}
\label{acknowledgements}

% Include this section in the table of contents
\addcontentsline{toc}{section}{Acknowledgements}
This work was supported by the Exascale Computing Project (17-SC-20-SC), a joint project of the U.S. Department of Energy’s Office of Science and National Nuclear Security Administration, responsible for delivering a capable exascale ecosystem, including software, applications, and hardware technology, to support the nation’s exascale computing imperative.

\end{document}