### 2024-04-11 RFCs/HDF5_Library/Float16/sections/acknowledgements.tex

\documentclass[../HDF5_RFC.tex]{subfiles}

\begin{document}

\section*{Acknowledgments}
\label{acknowledgements}

% Include this section in the table of contents
\addcontentsline{toc}{section}{Acknowledgements}

This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of
Fusion Energy Sciences, under Award Number DE-SC0024442.

\textbf{Disclaimer:} This report was prepared as an account of work sponsored by an agency of the United
States Government. Neither the United States Government nor any agency therefore, nor any of their employees,
makes any warranty, expressed or implied, or assumes any legal liability or responsibility for the accuracy,
completeness, or usefulness of any information, apparatus, product, or process disclosed, or represents that
this use would not infringe privately owned rights. Reference herein to any specific commercial product,
process, or service by trade name, trademark, manufacturer, or otherwise does not necessarily constitute or
imply its endorsement, recommendation, or favoring by the United States Government or any agency thereof.
The views and opinions of authors expressed herein do not necessarily state or reflect those of the United
States Government or any agency thereof.

\end{document}