### 2024-04-11 RFCs/HDF5_Library/Float16/sections/considerations.tex

\documentclass[../HDF5_RFC.tex]{subfiles}

\begin{document}

\section{Further considerations}
\label{considerations}

The following items are miscellaneous considerations to take into account when interacting with these
new datatypes:

\begin{itemize}
\item As there is no \texttt{printf} format specifier for the \texttt{\_Float16} C type, and since
a variable of type \texttt{\_Float16} will not be promoted to a larger type when passed to \texttt{printf},
one should cast a variable of this type to \texttt{double} and use the \texttt{\%f} specifier to print
the value of such a variable.

\end{itemize}

\end{document}