### 2024-04-11 RFCs/HDF5_Library/Float16/sections/tools.tex

\documentclass[../HDF5_RFC.tex]{subfiles}

\begin{document}

\section{HDF5 tools support}
\label{tools}

After adding these new datatypes, HDF5's tools library and tool applications will need to be
updated to understand these new types. At minimum, this will involve updating the tools library's
object dumping routines in \texttt{h5tools\_dump.c}, the \texttt{print\_type} routine in
\texttt{h5diff\_util.c} and the \texttt{print\_native\_type} routine in \texttt{h5ls.c}.
Additionally, tools tests which exercise using the tool applications on these new types should be
added to the library's existing tests.

As part of implementing support for these datatypes, it must be decided how the tool applications
should print out a visual representation for complex datatypes when encountered. This RFC proposes
the following as a starting point for discussion:

\begin{verbatim}
COMPLEX { real: VALUE, imag: VALUE }    
\end{verbatim}

\end{document}