### 2024-04-11 RFCs/HDF5_Library/Float16/sections/wrappers.tex

\documentclass[../HDF5_RFC.tex]{subfiles}

\begin{document}

\section{HDF5 language wrappers support}
\label{wrappers}

The following sections briefly detail implementing support for these new datatypes in the HDF5
language wrappers both internal and external to HDF5.

\subsection{Fortran}

\subsubsection{\texttt{\_Float16}}

Fortran support for this type appears to rely on compiler support; adding support in HDF5's Fortran
wrappers may be straightforward or may take a bit of work, depending on compiler support across platforms.

\subsubsection{\texttt{\_\_bf16}}

Fortran support for this type appears to rely on compiler support; adding support in HDF5's Fortran
wrappers may be straightforward or may take a bit of work, depending on compiler support across platforms.

\subsubsection{\texttt{\_Bool}}

Fortran has the \texttt{C\_BOOL} type to support this datatype.

\subsubsection{\texttt{float \_Complex}, \texttt{double \_Complex}, \texttt{long double \_Complex}}

Fortran has native support for this datatype; adding support in HDF5's Fortran wrappers should be straightforward.

\subsection{Python}

\subsubsection{\texttt{\_Float16}}

\texttt{h5py} supports this datatype through NumPY.

\subsubsection{\texttt{\_\_bf16}}

NumPY does not currently appear to support this type, but there has been some discussion around
supporting it, at which point \texttt{h5py} could support it as well.

\subsubsection{\texttt{\_Bool}}

\texttt{h5py} supports this datatype through NumPY.

\subsubsection{\texttt{float \_Complex}, \texttt{double \_Complex}, \texttt{long double \_Complex}}

\texttt{h5py} supports this datatype through NumPY.

\subsection{C++}

\subsubsection{\texttt{\_Float16}}

C++ has support for this type as an optional feature; adding support in HDF5's C++ wrappers may be
straightforward or may take a bit of work, depending on compiler support across platforms.

\subsubsection{\texttt{\_\_bf16}}

C++ has support for this type as an optional feature; adding support in HDF5's C++ wrappers may be
straightforward or may take a bit of work, depending on compiler support across platforms.

\subsubsection{\texttt{\_Bool}}

C++ has native support for this datatype; adding support in HDF5's C++ wrappers should be straightforward.

\subsubsection{\texttt{float \_Complex}, \texttt{double \_Complex}, \texttt{long double \_Complex}}

C++ has native support for these datatypes; adding support in HDF5's C++ wrappers should be straightforward.

\subsection{Java}

\subsubsection{\texttt{\_Float16}}

Java JDK 20 appears to introduce support for half-precision floating-point types, though they
seem to represent these values in the \texttt{short} datatype. For a Java JDK before this version,
support for this type could be implemented in HDF5's Java wrappers using one of several different
approaches.

\subsubsection{\texttt{\_\_bf16}}

Java doesn't support this datatype, but it could be implemented manually in HDF5's Java wrappers.

\subsubsection{\texttt{\_Bool}}

Java has native support for this datatype; adding support in HDF5's Java wrappers should be straightforward.

\subsubsection{\texttt{float \_Complex}, \texttt{double \_Complex}, \texttt{long double \_Complex}}

Support for complex numbers in HDF5's Java wrappers could be implemented through a Class.

\end{document}