### 2024-04-26 RFCs/HDF5_Library/WriteAheadLog/sections/abstract.tex
Add Write-Ahead Log RFC (version 1)

\documentclass[../HDF5_RFC.tex]{subfiles}

\begin{document}

\begin{abstract}
This RFC describes a new module that implements a Write-Ahead Log (WAL) for metadata within the HDF5 library. In the event of a crash during the lifetime of an application using HDF5, this WAL can be used by the library or an external tool to restore the metadata within an HDF5 file to a self-consistent state, preventing file corruption and limiting the loss of application data with minimal I/O and memory overhead.
\end{abstract}

\end{document}